---
title: "**Guns or Gandhi: Ethnic Elite Composition and Dissident Tactics in Civil Conflicts**"
author: 
  - "Ruslan Guseinov"
  - "Maaike Arienne Zeelen"
output:
    bookdown::pdf_document2:
    number_sections: false
    toc: false
linestretch: 2
indent: true
fontsize: 12pt
geometry: left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm
date: "`r Sys.Date()`"
bibliography: references.bib
header-includes:
  - \setcounter{secnumdepth}{0}
  - \usepackage{amsthm}
  - \newtheorem{hyp}{Hypothesis}
abstract: |
  This paper is very cool blbablalblab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,    # hide code from output
  warning = TRUE,    # suppress warnings
  message = FALSE,     # suppress package loading messages
  include = FALSE # hide code
)

stargazer_opt <- ifelse(knitr::is_latex_output(), "latex", "html")

### Set up ###------------------------------------------------------------------
## Clear environment
rm(list = ls())

## Load packages; Install packages if necessary
p_needed <- c(
  "tidyverse",        # dplyr, ggplot2, tidyr throughout
  "lme4",             # glmer for RE logit
  "MASS",             # mvrnorm in OVA simulation functions
  "kableExtra",       # kbl, kable_styling for all tables
  "patchwork",        # combining histogram + density plot
  "bookdown",         # pdf_document2, cross-referencing
  "countrycode",      # datacollection.R country coding
  "pROC",             # AUC-ROC model fit
  'modelsummary',     # For summary tables
  "car",              # For model fit
  "DHARMa",
  "splines"
)

packages <- rownames(installed.packages())
p_to_install <- p_needed[!(p_needed %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install) 
}

sapply(p_needed, require, character.only = TRUE)

## Load functions
source('functions/ova_sim_functions.R')    # use observed value approach (OVA) simulation functions

## Load dataset
df <- readRDS('datasources/df_EthnicElite_DissidentTactics.rds')
```

\section{Data and Methods}
\subsection{Data description}
We test our hypotheses using the NAVCO 2.1 dataset, which documents maximalist anti-government campaigns from 1945 to 2013 [@chenoweth2011civil]. The unit of analysis is the campaign onset, as our theory concerns the initial choice between violent and nonviolent resistance. Following NAVCO coding protocols, campaigns are categorized by their main tactic into two mutually exclusive groups: violent and nonviolent. We operationalize this distinction as a binary variable (1 = violent, 0 = nonviolent). Our final sample includes X violent and Y nonviolent campaign onsets.


Our key independent variables come from the Ethnic Power Relations (EPR) Core Dataset [@vogt2015integrating], which records politically relevant ethnic groups and their access to state power.

For Hypothesis 1, we follow Hendrix and Salehyan [-@hendrix2019ethnicity] and measure ethnic elite fragmentation using the Included Fractionalization Index (IFI):
$$
IFI = 1 - \sum_{i=1}^{N} s_i^2
$$
where (s_i) is the population share of included ethnic group (i). IFI captures the fragmentation of the ruling coalition. Unlike a simple count of included groups, it also accounts for their relative size. Higher values indicate more fragmented ruling coalitions, while lower values indicate that executive power is concentrated among fewer ethnic groups. 


For Hypothesis 2, we use a binary indicator derived from the highest EPR access-to-power category to distinguish between concentrated and shared executive power ([@Vogtetal2015]; [@HlatkyLandry2025]). For each campaign, the variable is coded as 1 if the highest status held by any politically relevant ethnic group is Senior Partner, indicating a meaningful power-sharing arrangement where executive authority is shared among multiple groups. It is coded as 0 if the highest status is Monopoly (exclusive control by a single group) or Dominant (one group controls the executive while others hold subordinate positions).


To reduce confounding, we control for standard economic, demographic, and political factors that may be correlated with both ethnic elite structures and campaign tactics. Economic controls include GDP per capita, GDP growth, and logged population size. Demographic controls include youth bulge, measured as the share of the population aged 15–29. Political controls include executive corruption, electoral democracy (polyarchy), regime duration, and peace years since the last maximalist campaign.

We also include excluded ethnic fractionalization (EFI), constructed analogously to IFI but using the population shares of excluded ethnic groups rather than included groups. This controls for ethnic heterogeneity among excluded populations, ensuring that effects attributed to ruling coalition fragmentation are not confounded by broader ethnic structure. Finally, we control for whether a campaign is initiated by an included ethnic group to ensure that results are not driven by systematic differences between campaigns initiated by included versus excluded groups.


Our theory focuses on pre-onset strategic decisions about campaign tactics. We therefore use lagged values of all independent variables from the previous year to capture the political environment in which these decisions are made.

Table @ref(tab:descriptive-stats) presents descriptive statistics for all variables, separately for violent and nonviolent campaigns. In the raw means, violent campaigns exhibit slightly higher levels of ethnic elite fragmentation (IFI) and more inclusive ethnic power-sharing arrangements within the executive compared to nonviolent campaigns. 

Missing data is limited and does not pose a systematic concern for the analysis. In total, only r round(sum(is.na(df_final_methods$epr_IFI_l)) / nrow(df_final_methods) * 100, 2)% of IFI values and r round(sum(is.na(df_final_methods$epr_highest_rank)) / nrow(df_final_methods) * 100, 2)% of EPR highest access-to-power observations are missing.

```{r descriptive-statistics}
## Create dataframe with categorized tactics and selected variables
summarydata <- df %>%
  mutate(
    Tactics = case_when(
      onset_type == "Nonviolent" ~ "Nonviolent",
      TRUE                       ~ "Violent"
    )
  ) %>%
  dplyr::select(
    Tactics,
    # Main independent variable for Hypothesis 1
    "Included Fractionalization Index" = epr_IFI_l,
    # Main independent variable for Hypothesis 2
    "Ethnic Power Inclusion Index"    = epr_highest_rank,
    
    # Control variables
    "Excluded Fractionalization Index"                = epr_EFI_l,
    "Ethnic Group in Power Participation"   = egip_participated_l,
    "ln(GDP per capita)" = log_gdp_pcap_l,
    "ln(Population)"     = pop_log_l,
    "Polyarchy"          = v2x_polyarchy_l,
    "Executive Corruption" = v2x_execorr_l,
    "ln(Regime Duration)"  = log_v2regdur_l,
    "Peace Years"          = peace_years_l
  )

## Generate table comparing means across violent and non-violent tactics
datasummary_balance(
  ~Tactics,
  data    = summarydata,
  title   = "Descriptive Statistics by Campaign Tactics (NAVCO 2.1 Onsets)",
  label   = "descriptive-stats",
  output  = "kableExtra",
  fmt     = 2,
  dinm    = FALSE
) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  
  ## Create table footnotes
  footnote(
    general = paste(
      "\nAll variables are one-year lagged.",
      "\nIncluded Fractionalization Index (IFI): Higher values indicate greater fragmentation among included ethnic groups.",
      "\nEthnic Power Fragmentation Index (EPFI): Higher values indicate more fragmented power distribution among included groups."
    ),
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  ) %>%
  
  ## Create section header for main variables
  pack_rows(
    "Main Variables",
    start_row = 1,
    end_row = 2,
    bold = TRUE
  ) %>%
  
  ## Create section header for control variables
  pack_rows(
    "Control Variables",
    start_row = 3,
    end_row = 10,
    bold = TRUE
  ) 
```





\subsection*{Modeling approach}
```{r variation-IV}
### Test variation of the independent variables ###-----------------------------
## Create function for measuring variation
# Define a function to calculate within-group, between-group, and total variation
# Returns the Intraclass Correlation Coefficient (ICC) to assess clustering strength

variance_decomp <- function(data, var, group) {
  
  ## Calculate group-level statistics
  data_grouped <- data %>%
    group_by(.data[[group]]) %>%
    mutate(
      group_mean = mean(.data[[var]], na.rm = TRUE),
      deviation = .data[[var]] - group_mean
    ) %>%
    ungroup()
  
  ## Calculate Total Variation
  overall_sd <- sd(data_grouped[[var]], na.rm = TRUE)
  
  ## Calculate Within-Group Variation
  within_sd <- sd(data_grouped$deviation, na.rm = TRUE)
  
  ## Calculate Between-Group Variation
  between_sd <- data_grouped %>%
    distinct(.data[[group]], group_mean) %>%
    summarise(
      sd = sd(group_mean, na.rm = TRUE)
    ) %>%
    pull(sd)
  
  ## Calculate Intraclass Correlation Coefficient (ICC)
  # Indicates proportion of total variance explained by group membership
  icc <- between_sd^2 / (between_sd^2 + within_sd^2)
  
  ## Return results as a tidy tibble
  tibble(
    variable = var,
    grouping = group,
    overall_sd = overall_sd,
    within_sd = within_sd,
    between_sd = between_sd,
    ICC = icc
  )
}

## Define key independent variables
vars <- c("epr_IFI_l", "epr_highest_rank")

## Decompose variance by country
variance_results_country <- map_dfr(
  vars,
  variance_decomp,
  data = df,
  group = "cow"
)

## Clean environment
rm(list = c("summarydata", "variance_results_country", "variance_decomp", "vars"))
```

Because the dependent variable captures whether a campaign onset is violent or nonviolent, we estimate logistic regression models. The focus of the analysis is on the choice of tactic at campaign onset, rather than the timing or duration of campaigns; a binary logit specification is therefore appropriate.

To account for the panel structure of the data—where multiple campaign onsets may occur within the same country—we include country-level varying intercepts. This allows baseline differences in the propensity toward violent versus nonviolent campaign onset to vary across countries.

Ethnic configurations change slowly over time, and exploratory analysis indicates limited within-country variation in both IFI and EFI. As a result, a fixed-effects specification would rely primarily on the relatively small subset of countries that experience meaningful changes in ethnic inclusion structures across campaign onsets. Given that our theoretical argument concerns how both cross-national differences and temporal variation in elite ethnic structures shape strategic choices, we retain both sources of variation. A random-intercepts specification is therefore more consistent with the structure of the data and the substantive scope of the hypotheses.

We acknowledge, however, that random-effects models rely on a stronger assumption than fixed effects—namely that country-specific unobserved heterogeneity is uncorrelated with the included regressors. If this assumption is violated, coefficient estimates may be biased [@clark2015should]. Despite this limitation, the random-intercepts approach represents an appropriate trade-off between bias and efficiency in this setting, given the limited within-country variation in key predictors.

For Hypothesis 1, our model specification is:
$$
\begin{aligned}
& \text{violent}_{it} \sim \text{Bernoulli}(\pi_{it}) \\[0.3em]
& \log\left(\frac{\pi_{it}}{1-\pi_{it}}\right) = \beta_0 
+ \beta_1 (\text{IFI})_{it-1} 
+ \beta_2 (\text{EFI})_{it-1} 
+ \mathbf{X}'_{it-1}\boldsymbol{\gamma} + \alpha_{[j]i} \\[0.3em]
& \alpha_j \sim \mathcal{N}(\mu_\alpha, \sigma^2_\alpha)
\end{aligned}
$$
where $\mathbf{X}_{it-1}$ denotes the vector of lagged control variables and $u_i$ is a country-specific random intercept capturing unobserved time-invariant heterogeneity.


\section{Results}
```{r H1-analysis}
### H1: Analysis ###------------------------------------------------------------
## Define a formula for H1
formula_h1 <- NVC2.1_VIOL ~ epr_IFI_l + epr_EFI_l + egip_participated_l +
  log_gdp_pcap_l + pop_log_l + v2x_polyarchy_l + v2x_execorr_l +
  log_v2regdur_l + peace_years_l + (1 | cow)  # random intercept by country

model1 <- glmer(formula_h1,
                data = df,
                family = binomial(link = "logit"),
                control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) # apply an optimizer for convergence

## Simulate predicted probabilities (Observed value approach)
simulation_number = 1000
ova_egip_full_m1  <- sim_ova_range(model1,        
                                   focal_var = "epr_IFI_l", 
                                   nsim = simulation_number,
                                   vcov_matrix = vcov(model1))$result %>% mutate(type = "Full Sample")

## Create plot for predicted probabilities 
ggplot(ova_egip_full_m1,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df$NVC2.1_VIOL) / nrow(df), 2), # a benchmark probability dashed line
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data = df %>% drop_na(epr_IFI_l),
    aes(x = epr_IFI_l),
    inherit.aes = FALSE,
    color = "#C0392B", alpha = 0.3,
    sides = "b"          # bottom only
  ) + 
  labs(x     = "IFI (0 = fully concentrated, 1 = fully fragmented)", 
       y     = "Predicted Probability \n Violent Campaign", 
       caption = paste0(
         simulation_number, " simulations, 95% CI\n",
         "Note: The dashed line represents the empirical probability of a violent event in the estimation sample (i.e., the mean of the binary outcome)."
       )) +
  theme_bw() +
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```
```{r H2-analysis}
### H2: Analysis ###------------------------------------------------------------
## Define a formula for H2
formula_h2 <- NVC2.1_VIOL ~ epr_highest_rank + epr_EFI_l + egip_participated_l +
  log_gdp_pcap_l + pop_log_l + v2x_polyarchy_l + v2x_execorr_l +
  log_v2regdur_l + peace_years_l + (1 | cow)  # random intercept by country

model2 <- glmer(formula_h2,
                data = df,
                family = binomial(link = "logit"),
                control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) # apply an optimizer for convergence

#WARNING: A convergence issue occured. Reestiamte a model with a warm start
all_fits_m2 <- allFit(model2)
ok_fits  <- all_fits_m2[summary(all_fits_m2)$which.OK]
llik_vals <- sapply(ok_fits, logLik)
model2  <- ok_fits[[which.max(llik_vals)]]
model2@optinfo$conv$lme4 #model converged

## Simulate predicted probabilities (Observed value approach)
ova_egip_full_m2  <- sim_ova_range(model2,        
                                   focal_var = "epr_highest_rank", 
                                   nsim = simulation_number,
                                   vcov_matrix = vcov(model2))$result %>% mutate(type = "Full Sample")

## Create plot for predicted probabilities 
ggplot(ova_egip_full_m2,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df$NVC2.1_VIOL) / nrow(df), 2), # a benchmark probability dashed line
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data = df %>% drop_na(epr_highest_rank),
    aes(x = epr_highest_rank),
    inherit.aes = FALSE,
    color = "#C0392B", alpha = 0.3,
    sides = "b"          # bottom only
  ) + 
  labs(x     = "Power Equality Index", 
       y     = "Predicted Probability \n Violent Campaign", 
       caption = paste0(
         simulation_number, " simulations, 95% CI\n",
         "Note: The dashed line represents the empirical probability of a violent event in the estimation sample (i.e., the mean of the binary outcome)."
       )) +
  theme_bw() +
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```

\section{Model fit}
```{r H1-confusion-matrix}
### Create confusion matrix ###
## Set same seed as simulations for reproducibility
set.seed(1234)

## Define the number of observations
n_total <- nrow(df)

## Randomly select 2/3 of observations for training
train_idx <- sample(1:n_total, size = floor(2/3 * n_total), replace = FALSE)

## Select remaining 1/3 used as test set
test_idx  <- setdiff(1:n_total, train_idx)

## Construct training dataset
df_train <- df[train_idx, ]

## Estimate mixed-effects logistic regression on training data
model1_train  <- glmer(formula_h1,        # random intercept by country
                       data = df_train,
                       family = binomial(link = "logit"),
                       control = glmerControl(optimizer = "bobyqa",
                                              optCtrl = list(maxfun = 2e5)))

## Extract countries that are present in the training sample
trained_cows <- rownames(ranef(model1_train)$cow)

## Construct test dataset
df_test <- df[test_idx, ] %>%
  filter(as.character(cow) %in% trained_cows) %>%  # test is only with the trained countries
  drop_na(epr_IFI_l, epr_EFI_l, egip_participated_l, log_gdp_pcap_l, 
          pop_log_l, v2x_execorr_l, v2x_polyarchy_l, log_v2regdur_l, 
          peace_years_l) # remove observations with missing values in model covariates

## Predict probabilities of violence for test data
pred_test_prob <- predict(model1_train, newdata = df_test, type = "response")

## Define actual outcome labels
actual_test <- ifelse(df_test$NVC2.1_VIOL == 1,
                      "Violent", "Nonviolent")

## Classify predictions using a 0.5 probability threshold
pred_test   <- ifelse(pred_test_prob >= 0.5, "Violent", "Nonviolent")

## Compute ROC curve and AUC
roc_obj  <- roc(response  = ifelse(actual_test == "Violent", 1, 0),
                predictor = pred_test_prob,
                quiet     = TRUE)
auc_val  <- as.numeric(auc(roc_obj))

## Construct confusion matrix components
tp <- sum(pred_test == "Violent"    & actual_test == "Violent")
tn <- sum(pred_test == "Nonviolent" & actual_test == "Nonviolent")
fp <- sum(pred_test == "Violent"    & actual_test == "Nonviolent")
fn <- sum(pred_test == "Nonviolent" & actual_test == "Violent")

## Calculate metrics
accuracy  <- (tp + tn) / (tp + tn + fp + fn)
precision <- tp / (tp + fp)
recall    <- tp / (tp + fn)   
f1        <- 2 * (precision * recall) / (precision + recall)

## Create confusion matrix table
conf_df <- data.frame(
  ` `            = c("Predicted Nonviolent", "Predicted Violent"),
  `Actual Nonviolent` = c(as.character(tn), as.character(fp)),
  `Actual Violent`    = c(as.character(fn), as.character(tp)),
  check.names = FALSE
)

## Create performance metrics table
metrics_df <- data.frame(
  ` `                 = c("Accuracy", "Precision", 
                          "Recall",  "F1 Score", "AUC"),
  `Actual Nonviolent` = c(sprintf("%.3f", accuracy), 
                          sprintf("%.3f", precision),
                          sprintf("%.3f", recall), 
                          sprintf("%.3f", f1),
                          sprintf("%.3f", auc_val)),
  `Actual Violent`    = c("", "", "", "", ""),
  check.names = FALSE
)

## Combine confusion matrix and metrics
bind_rows(conf_df, metrics_df) %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Confusion Matrix and Predictive Performance — Model 1",
    label    = "conf-matrix-h1",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(2, extra_css = "border-bottom: 2px solid black;") %>%  # separator line
  pack_rows("Confusion Matrix", 1, 2) %>%
  pack_rows("Performance Metrics", 3, 7)
```






```{r H2-confusion-matrix}
### Create confusion matrix ###
## Set same seed as simulations for reproducibility
set.seed(1234)

## Define the number of observations
n_total <- nrow(df)

## Randomly select 2/3 of observations for training
train_idx <- sample(1:n_total, size = floor(2/3 * n_total), replace = FALSE)

## Select remaining 1/3 used as test set
test_idx  <- setdiff(1:n_total, train_idx)

## Construct training dataset
df_train <- df[train_idx, ]

## Estimate mixed-effects logistic regression on training data
model2_train  <- glmer(formula_h2,        # random intercept by country
                       data = df_train,
                       family = binomial(link = "logit"),
                       control = glmerControl(optimizer = "bobyqa",
                                              optCtrl = list(maxfun = 2e5)))

## Extract countries that are present in the training sample
trained_cows <- rownames(ranef(model2_train)$cow)

## Construct test dataset
df_test <- df[test_idx, ] %>%
  filter(as.character(cow) %in% trained_cows) %>%  # test is only with the trained countries
  drop_na(epr_highest_rank, egip_participated_l, log_gdp_pcap_l, 
          pop_log_l, v2x_execorr_l, v2x_polyarchy_l, log_v2regdur_l, 
          peace_years_l) # remove observations with missing values in model covariates

## Predict probabilities of violence for test data
pred_test_prob <- predict(model2_train, newdata = df_test, type = "response")

## Define actual outcome labels
actual_test <- ifelse(df_test$NVC2.1_VIOL == 1,
                      "Violent", "Nonviolent")

## Classify predictions using a 0.5 probability threshold
pred_test   <- ifelse(pred_test_prob >= 0.5, "Violent", "Nonviolent")

## Compute ROC curve and AUC
roc_obj  <- roc(response  = ifelse(actual_test == "Violent", 1, 0),
                predictor = pred_test_prob,
                quiet     = TRUE)
auc_val  <- as.numeric(auc(roc_obj))

## Construct confusion matrix components
tp <- sum(pred_test == "Violent"    & actual_test == "Violent")
tn <- sum(pred_test == "Nonviolent" & actual_test == "Nonviolent")
fp <- sum(pred_test == "Violent"    & actual_test == "Nonviolent")
fn <- sum(pred_test == "Nonviolent" & actual_test == "Violent")

## Calculate metrics
accuracy  <- (tp + tn) / (tp + tn + fp + fn)
precision <- tp / (tp + fp)
recall    <- tp / (tp + fn)   
f1        <- 2 * (precision * recall) / (precision + recall)

## Create confusion matrix table
conf_df <- data.frame(
  ` `            = c("Predicted Nonviolent", "Predicted Violent"),
  `Actual Nonviolent` = c(as.character(tn), as.character(fp)),
  `Actual Violent`    = c(as.character(fn), as.character(tp)),
  check.names = FALSE
)

## Create performance metrics table
metrics_df <- data.frame(
  ` `                 = c("Accuracy", "Precision", 
                          "Recall",  "F1 Score", "AUC"),
  `Actual Nonviolent` = c(sprintf("%.3f", accuracy), 
                          sprintf("%.3f", precision),
                          sprintf("%.3f", recall), 
                          sprintf("%.3f", f1),
                          sprintf("%.3f", auc_val)),
  `Actual Violent`    = c("", "", "", "", ""),
  check.names = FALSE
)

## Combine confusion matrix and metrics
bind_rows(conf_df, metrics_df) %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Confusion Matrix and Predictive Performance — Model 1",
    label    = "conf-matrix-h1",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(2, extra_css = "border-bottom: 2px solid black;") %>%  # separator line
  pack_rows("Confusion Matrix", 1, 2) %>%
  pack_rows("Performance Metrics", 3, 7)
```


\section{Assumption Testing}







\section{Appendix Model Fit}


```{r H1-auc-roc-curve}
### H1: AUC-ROC curve ###-------------------------------------------------------
# Step 1: train-test split (same seed as confusion matrix section)
set.seed(1234)

pred_test <- predict(model1_train, newdata = df_test,
                     type = "response",
                     allow.new.levels = TRUE)

valid_idx    <- !is.na(pred_test) & !is.na(df_test$NVC2.1_VIOL)
actual_test  <- df_test$NVC2.1_VIOL[valid_idx]
probs_test   <- pred_test[valid_idx]

roc_obj <- roc(response  = actual_test,
               predictor = probs_test,
               smooth    = FALSE,      # kernel smoothing
               quiet     = TRUE)

roc_df <- data.frame(
  FPR = 1 - roc_obj$specificities,
  TPR = roc_obj$sensitivities
)

ggplot(roc_df, aes(x = FPR, y = TPR)) +
  geom_ribbon(aes(ymin = 0, ymax = TPR),
              fill = "#C0392B", alpha = 0.08) +
  geom_line(color = "#C0392B", linewidth = 0.9) +
  geom_abline(slope     = 1, intercept = 0,
              linetype  = "dashed", color = "grey60",
              linewidth = 0.5) +
  annotate("text",
           x     = 0.72, y = 0.25,
           label = paste0("AUC = ", round(auc(roc_obj), 3)),
           color = "#C0392B", size = 4) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  labs(
    x       = "False positive rate (1 - Specificity)",
    y       = "True positive rate (Sensitivity)"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(size = 8, color = "grey50", hjust = 1)
  )

```
```{r H2-auc-roc-curve}
### H2: AUC-ROC curve ###-------------------------------------------------------
set.seed(1234)

pred_test <- predict(model2_train, newdata = df_test,
                     type = "response",
                     allow.new.levels = TRUE)

valid_idx    <- !is.na(pred_test) & !is.na(df_test$NVC2.1_VIOL)
actual_test  <- df_test$NVC2.1_VIOL[valid_idx]
probs_test   <- pred_test[valid_idx]

roc_obj <- roc(response  = actual_test,
               predictor = probs_test,
               smooth    = FALSE,      
               quiet     = TRUE)

roc_df <- data.frame(
  FPR = 1 - roc_obj$specificities,
  TPR = roc_obj$sensitivities
)

ggplot(roc_df, aes(x = FPR, y = TPR)) +
  geom_ribbon(aes(ymin = 0, ymax = TPR),
              fill = "#C0392B", alpha = 0.08) +
  geom_line(color = "#C0392B", linewidth = 0.9) +
  geom_abline(slope     = 1, intercept = 0,
              linetype  = "dashed", color = "grey60",
              linewidth = 0.5) +
  annotate("text",
           x     = 0.72, y = 0.25,
           label = paste0("AUC = ", round(auc(roc_obj), 3)),
           color = "#C0392B", size = 4) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  labs(
    x       = "False positive rate (1 - Specificity)",
    y       = "True positive rate (Sensitivity)"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(size = 8, color = "grey50", hjust = 1)
  )
```

```{r cross-validation-function}
### Function for cross validation ###-------------------------------------------
cross_validation_fun_glmer <- function(validation_number = 50,
                                       data,
                                       dep_var,
                                       main_iv) {
  
  formula_obj <- as.formula(
    paste(
      dep_var,
      "~",
      main_iv,
      "+ epr_EFI_l + log_gdp_pcap_l +",
      "egip_participated_l + pop_log_l +",
      "v2x_polyarchy_l + v2x_execorr_l +",
      "log_v2regdur_l + peace_years_l +",
      "(1 | cow)"
    ))
  
  data_realizations    <- data[data[[dep_var]] == 1, ]
  data_nonrealizations <- data[data[[dep_var]] == 0, ]
  
  test_fit   <- glmer(formula_obj, family = binomial("logit"), data = data,
                      control = glmerControl(optimizer = "bobyqa",
                                             optCtrl   = list(maxfun = 2e5)))
  coef_names <- names(fixef(test_fit))
  n_coefs    <- length(coef_names)
  
  # pre-allocate for target number of CONVERGED models
  coefs_matrix <- matrix(NA, nrow = validation_number, ncol = n_coefs)
  var_matrix   <- matrix(NA, nrow = validation_number, ncol = n_coefs)
  colnames(coefs_matrix) <- coef_names
  colnames(var_matrix)   <- coef_names
  
  n_converged  <- 0                        # converged iterations collected
  n_attempts   <- 0                        # total attempts made
  max_attempts <- validation_number * 10   # stop if too many failures
  current_seed <- 1
  
  while (n_converged < validation_number && n_attempts < max_attempts) {
    
    n_attempts   <- n_attempts + 1
    set.seed(current_seed)
    current_seed <- current_seed + 1
    
    samp1 <- sample(1:nrow(data_realizations),
                    size = floor(2/3 * nrow(data_realizations)))
    samp0 <- sample(1:nrow(data_nonrealizations),
                    size = floor(2/3 * nrow(data_nonrealizations)))
    sample_data <- rbind(data_realizations[samp1, ],
                         data_nonrealizations[samp0, ])
    
    
    model <- glmer(formula_obj, family = binomial("logit"), data = sample_data,
                   control = glmerControl(optimizer = "bobyqa",
                                          optCtrl = list(maxfun = 2e5)))
    
    #check convergence
    if (!is.null(model@optinfo$conv$lme4$messages)) {
      n_converged <- n_converged + 1
      
      vec_coefs <- fixef(model)
      vec_var   <- diag(vcov(model))
      idx       <- match(names(vec_coefs), coef_names)
      idx_valid <- !is.na(idx)
      coefs_matrix[n_converged, idx[idx_valid]] <- vec_coefs[idx_valid]
      var_matrix[n_converged,   idx[idx_valid]] <- vec_var[idx_valid]
      
      cat("Attempt", n_attempts, "| Converged:", n_converged, "/",
          validation_number, "\n")
    } else {
      cat("Attempt", n_attempts, "| Did not converge — skipping\n")
    }
  }
  
  # report final status
  if (n_converged < validation_number) {
    warning(paste0("Only ", n_converged, " of ", validation_number,
                   " converged models collected after ", 
                   max_attempts, " attempts."))
  }
  
  # trim matrices to actual converged rows
  coefs_matrix <- coefs_matrix[1:n_converged, , drop = FALSE]
  var_matrix   <- var_matrix[1:n_converged,   , drop = FALSE]
  
  # King (2001) combination rule — use n_converged throughout
  mean_coef   <- colMeans(coefs_matrix, na.rm = TRUE)
  W           <- colMeans(var_matrix,   na.rm = TRUE)
  B           <- apply(coefs_matrix, 2, var, na.rm = TRUE)
  T_var       <- W + (1 + 1 / n_converged) * B   # n_converged not n
  se_combined <- sqrt(T_var)
  z_value     <- mean_coef / se_combined
  p_value     <- 2 * (1 - pnorm(abs(z_value)))
  stars <- ifelse(p_value < 0.001, "***",
                  ifelse(p_value < 0.01,  "**",
                         ifelse(p_value < 0.05,  "*",
                                ifelse(p_value < 0.1,   ".", ""))))
  
  results <- data.frame(
    Variable     = coef_names,
    Coefficient  = round(mean_coef,   6),
    `Std. Error` = round(se_combined, 6),
    `z value`    = round(z_value,     3),
    `Pr(>|z|)`   = round(p_value,     4),
    Sig          = stars,
    check.names  = FALSE
  )
  
  return(list(
    results        = results,
    combined_coefs = mean_coef,
    combined_se    = se_combined,
    coefs_matrix   = coefs_matrix,
    var_matrix     = var_matrix,
    n_converged    = n_converged,
    n_attempts     = n_attempts
  ))
}
```

```{r cross-validation}
### H1: Apply cross validation function ###-------------------------------------
val_num <- 5
cross_val <- cross_validation_fun_glmer(data = df, dep_var = 'NVC2.1_VIOL', validation_number = val_num, 
                                        main_iv = "epr_IFI_l")


# Extract main model coefficients and SEs
main_coefs <- fixef(model1)
main_se    <- sqrt(diag(vcov(model1)))

# Cross-validated results
cv_coefs <- cross_val$combined_coefs
cv_se    <- cross_val$combined_se

var_labels <- c(
  "epr_IFI_l"         = "IFI",
  "epr_EFI_l"         = "EFI",
  "egip_participated_l" = 'Ethnic Group in Power Participation',
  "log_gdp_pcap_l"               = "ln(GDP per capita)",
  "pop_log_l"                    = "ln(Population)",
  "v2x_execorr_l"                = "Executive Corruption",
  "v2x_polyarchy_l"              = "Polyarchy",
  "log_v2regdur_l"               = "ln(Regime Duration)",
  "peace_years_l"                = "Peace Years"
)

# Combine into table
comparison_df <- data.frame(
  Variable        = names(main_coefs),
  `Main Model`    = sprintf("%.3f (%.3f)", main_coefs, main_se),
  `Cross-Validated` = sprintf("%.3f (%.3f)", cv_coefs, cv_se),
  check.names     = FALSE
) %>%
  filter(Variable != "(Intercept)") %>%
  mutate(Variable = dplyr::recode(Variable, !!!var_labels))  # apply labels

comparison_df %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Model 1: Main vs. Cross-Validated Coefficients",
    label    = "cv-comparison",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(
    general = "Intercept is not reported. Standard errors in parentheses. Cross-validated SE via King (2001) combination rule.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable    = TRUE 
  )


### Apply cross validation function ###-------------------------------------

val_num <- 5

# Run cross-validation for both models
cv_model1 <- cross_validation_fun_glmer(data = df, dep_var = 'NVC2.1_VIOL', 
                                        validation_number = val_num, 
                                        main_iv = "epr_IFI_l")

cv_model2 <- cross_validation_fun_glmer(data = df, dep_var = 'NVC2.1_VIOL', 
                                        validation_number = val_num, 
                                        main_iv = "epr_highest_rank")

# Extract coefficients and SEs for both models
main_coefs1 <- fixef(model1)
main_se1    <- sqrt(diag(vcov(model1)))
cv_coefs1   <- cv_model1$combined_coefs
cv_se1      <- cv_model1$combined_se

main_coefs2 <- fixef(model2)
main_se2    <- sqrt(diag(vcov(model2)))
cv_coefs2   <- cv_model2$combined_coefs
cv_se2      <- cv_model2$combined_se

# Common variable labels
var_labels <- c(
  "epr_IFI_l"           = "IFI",
  "epr_highest_rank"    = "Ethnic Power Fragmentation Index",
  "epr_EFI_l"           = "EFI",
  "egip_participated_l" = "Ethnic Group in Power Participation",
  "log_gdp_pcap_l"      = "ln(GDP per capita)",
  "pop_log_l"           = "ln(Population)",
  "v2x_execorr_l"       = "Executive Corruption",
  "v2x_polyarchy_l"     = "Polyarchy",
  "log_v2regdur_l"      = "ln(Regime Duration)",
  "peace_years_l"       = "Peace Years"
)

# Create data frame for Model 1
df1 <- data.frame(
  Variable        = names(main_coefs1),
  `Model 1 Main`  = sprintf("%.3f (%.3f)", main_coefs1, main_se1),
  `Model 1 CV`    = sprintf("%.3f (%.3f)", cv_coefs1, cv_se1),
  check.names     = FALSE
) %>%
  filter(Variable != "(Intercept)") %>%
  mutate(Variable = dplyr::recode(Variable, !!!var_labels))

# Create data frame for Model 2
df2 <- data.frame(
  Variable        = names(main_coefs2),
  `Model 2 Main`  = sprintf("%.3f (%.3f)", main_coefs2, main_se2),
  `Model 2 CV`    = sprintf("%.3f (%.3f)", cv_coefs2, cv_se2),
  check.names     = FALSE
) %>%
  filter(Variable != "(Intercept)") %>%
  mutate(Variable = dplyr::recode(Variable, !!!var_labels))

# Combine both data frames
comparison_df <- full_join(df1, df2, by = "Variable")

# Create combined table
comparison_df %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c", "c", "c"),
    caption  = "Cross-Validation Comparison: Model 1 vs Model 2",
    label    = "cv-comparison-combined",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(
    general = "Intercept is not reported. Standard errors in parentheses. Cross-validated SE via King (2001) combination rule.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable    = TRUE 
  )
```


\section{Appendix Assumptions}
```{r check-multicollinearity}
## Check for multicollinearity
vif(model1)
vif(model2)
```

```{r residual-diagnostics}
## Plot residuals vs fitted
plot(fitted(model1), residuals(model1, type = "pearson"))
abline(h = 0)

plot(fitted(model2), residuals(model2, type = "pearson"))
abline(h = 0)
```

```{r DHARMa-diagnostics}
## Create QQ plot
sim1 <- simulateResiduals(model1)
plot(sim1)

testUniformity(sim1)
testDispersion(sim1)

## Create QQ plot
sim2 <- simulateResiduals(model2)
plot(sim2)

testUniformity(sim2)
testDispersion(sim2)
```

```{r splines-models}
## Check Linearity of continuous predictors on the logit scale
model1_splines <- glmer(NVC2.1_VIOL ~ ns(epr_IFI_l, df = 3) + ns(epr_EFI_l, df = 3)  + egip_participated_l +
                          log_gdp_pcap_l + pop_log_l + v2x_polyarchy_l + v2x_execorr_l +
                          log_v2regdur_l + peace_years_l + (1 | cow),
                        data = df,
                        family = binomial(link = "logit"),
                        control = glmerControl(optimizer = "bobyqa",
                                               optCtrl = list(maxfun = 2e5)))
#WARNING: A convergence issue occured. Reestiamte a model with a warm start
all_fits_m1splines <- allFit(model1_splines)
ok_fits  <- all_fits_m1splines[summary(all_fits_m1splines)$which.OK]
llik_vals <- sapply(ok_fits, logLik)
model1_splines  <- ok_fits[[which.max(llik_vals)]]

## Check Linearity of continuous predictors on the logit scale
model2_splines <- glmer(NVC2.1_VIOL ~ ns(epr_highest_rank, df = 3) + ns(epr_EFI_l, df = 3)  + egip_participated_l +
                          log_gdp_pcap_l + pop_log_l + v2x_polyarchy_l + v2x_execorr_l +
                          log_v2regdur_l + peace_years_l + (1 | cow),
                        data = df,
                        family = binomial(link = "logit"),
                        control = glmerControl(optimizer = "bobyqa",
                                               optCtrl = list(maxfun = 2e5)))
#WARNING: A convergence issue occured. Reestiamte a model with a warm start
all_fits_m2splines <- allFit(model2_splines)
ok_fits  <- all_fits_m2splines[summary(all_fits_m2splines)$which.OK]
llik_vals <- sapply(ok_fits, logLik)
model2_splines  <- ok_fits[[which.max(llik_vals)]]

#Models converged
#model1_splines@optinfo$conv$lme4
#model2_splines@optinfo$conv$lme4
```

```{r influence-diagnostics}
## Calculate cook's distance
cook_obs_m1 <- cooks.distance(influence(model1, obs=TRUE))

## Rerun model without influential outliers
model_obs_m1 <- glmer(formula_h1,
                data = df[cook_obs_m1 <= 4/length(cook_obs_m1), ],
                family = binomial(link = "logit"),
                control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) # apply an optimizer for convergence


## Calculate cook's distance
cook_obs_m2 <- cooks.distance(influence(model2, obs=TRUE))

## Rerun model without influential outliers
model_obs_m2 <- glmer(formula_h2,
                data = df[cook_obs_m2 <= 4/length(cook_obs_m2), ],
                family = binomial(link = "logit"),
                control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) # apply an optimizer for convergence
```


```{r random-effects-variance}
## Check variance of random effect
VarCorr(model1)

VarCorr(model2)

```

```{r h1-REWB-model}
### H1: Random Effects Within-Between (REWB) model ###--------------------------
ova_egip_full_m1_appendix <- ova_egip_full_m1 %>% mutate(type = "RE")

rewb_model <- glmer(NVC2.1_VIOL ~
                      demeaned_IFI + mean_IFI +
                      demeaned_EFI + mean_EFI +
                      log_gdp_pcap_l + pop_log_l +
                      v2x_execorr_l + v2x_polyarchy_l +
                      log_v2regdur_l + peace_years_l +
                      (1 | cow),
                    data   = df,
                    family = binomial("logit"))
#WARNING: A convergence issue occured. Reestiamte a model with a warm start
all_fits_rewb_model <- allFit(rewb_model)
ok_fits  <- all_fits_rewb_model[summary(all_fits_rewb_model)$which.OK]
llik_vals <- sapply(ok_fits, logLik)
rewb_model  <- ok_fits[[which.max(llik_vals)]]
#rewb_model@optinfo$conv$lme4 #model converged

ova_egip_full_rewb  <- sim_ova_range(rewb_model,        
                                     focal_var = "mean_IFI", 
                                     nsim = simulation_number,
                                     vcov_matrix = vcov(rewb_model))$result %>% mutate(type = "REWB")

ova_model1_robustness_appendix <- bind_rows(
  ova_egip_full_m1_appendix,
  ova_egip_full_rewb
) %>%
  mutate(type = factor(type, levels = c(
    "RE",
    "REWB"
  )))

rug_data <- bind_rows(
  df %>% drop_na(epr_IFI_l) %>%
    mutate(type = "RE", rug_x = epr_IFI_l),
  df %>% drop_na(mean_IFI) %>%
    mutate(type = "REWB", rug_x = mean_IFI)
) %>%
  mutate(type = factor(type, levels = c("RE", "REWB")))

ggplot(ova_model1_robustness_appendix,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df$NVC2.1_VIOL) / nrow(df), 2), 
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data        = rug_data,
    aes(x       = rug_x),
    inherit.aes = FALSE,
    color       = "#C0392B", alpha = 0.3,
    sides       = "b"
  ) + 
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```

```{r h2-REWB-model}
### H2: Random Effects Within-Between (REWB) model ###--------------------------
ova_egip_full_m2_appendix <- ova_egip_full_m2 %>% mutate(type = "RE")

rewb_model <- glmer(NVC2.1_VIOL ~
                      demeaned_epr_highest_rank +
                      mean_epr_highest_rank +
                      demeaned_EFI + mean_EFI +
                      log_gdp_pcap_l + pop_log_l +
                      v2x_execorr_l + v2x_polyarchy_l +
                      log_v2regdur_l + peace_years_l +
                      (1 | cow),
                    data   = df,
                    family = binomial("logit"))
#WARNING: A convergence issue occured. Reestiamte a model with a warm start
all_fits_rewb_model <- allFit(rewb_model)
ok_fits  <- all_fits_rewb_model[summary(all_fits_rewb_model)$which.OK]
llik_vals <- sapply(ok_fits, logLik)
rewb_model  <- ok_fits[[which.max(llik_vals)]]
#rewb_model@optinfo$conv$lme4 #model converged

ova_egip_full_rewb  <- sim_ova_range(rewb_model,        
                                     focal_var = "mean_epr_highest_rank", 
                                     nsim = simulation_number,
                                     vcov_matrix = vcov(rewb_model))$result %>% mutate(type = "REWB")

ova_model2_robustness_appendix <- bind_rows(
  ova_egip_full_m2_appendix,
  ova_egip_full_rewb
) %>%
  mutate(type = factor(type, levels = c(
    "RE",
    "REWB"
  )))

rug_data <- bind_rows(
  df %>% drop_na(epr_highest_rank) %>%
    mutate(type = "RE", rug_x = epr_highest_rank),
  df %>% drop_na(mean_IFI) %>%
    mutate(type = "REWB", rug_x = epr_highest_rank)
) %>%
  mutate(type = factor(type, levels = c("RE", "REWB")))

ggplot(ova_model1_robustness_appendix,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df$NVC2.1_VIOL) / nrow(df), 2), 
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data        = rug_data,
    aes(x       = rug_x),
    inherit.aes = FALSE,
    color       = "#C0392B", alpha = 0.3,
    sides       = "b"
  ) + 
  labs(x     = "", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```









\section{Appendix Robustness Checks}
```{r h1-subsample-models}
### H1: Robustness Checks ###---------------------------------------------------

model1_africa <- glmer(formula_h1,
                       family = binomial(link = "logit"),
                       data = df %>% filter(region %in% c('Middle East & North Africa', 'Sub-Saharan Africa')),
                       control = glmerControl(optimizer = "bobyqa",
                                              optCtrl = list(maxfun = 2e5))) # filter by the regionss

model1_noterr <- glmer(formula_h1,
                       family = binomial(link = "logit"),
                       data = df %>% filter(NVC2.1_territorial == 0),
                       control = glmerControl(optimizer = "bobyqa",
                                              optCtrl = list(maxfun = 2e5))) #filter by the non-territorial campaigns


ova_egip_africa_m1  <- sim_ova_range(model1_africa,        
                                     focal_var = "epr_IFI_l", 
                                     nsim = simulation_number,
                                     vcov_matrix = vcov(model1_africa))$result %>% mutate(type = "MENA & Sub-Saharan Africa")
ova_egip_noterr_m1  <- sim_ova_range(model1_noterr,        
                                     focal_var = "epr_IFI_l", 
                                     nsim = simulation_number,
                                     vcov_matrix = vcov(model1_noterr))$result %>% mutate(type = "Excl. territorial campaigns")

ova_model1_robustness <- bind_rows(
  ova_egip_africa_m1,
  ova_egip_noterr_m1,
) %>%
  mutate(type = factor(type, levels = c(
    "MENA & Sub-Saharan Africa",
    "Excl. territorial campaigns"
  )))


ggplot(ova_model1_robustness,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df$NVC2.1_VIOL) / nrow(df), 2),  # a benchmark probability dashed line
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data = df %>% drop_na(epr_IFI_l),
    aes(x = epr_IFI_l),
    inherit.aes = FALSE,
    color = "#C0392B", alpha = 0.3,
    sides = "b"          # bottom only
  ) + 
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )

```

```{r h2-subsample-models}
### H2: Robustness Checks ###---------------------------------------------------

model2_africa <- glmer(formula_h2,
                       family = binomial(link = "logit"),
                       data = df %>% filter(region %in% c('Middle East & North Africa', 'Sub-Saharan Africa')),
                       control = glmerControl(optimizer = "bobyqa",
                                              optCtrl = list(maxfun = 2e5))) # filter by the regionss

model2_noterr <- glmer(formula_h2,
                       family = binomial(link = "logit"),
                       data = df %>% filter(NVC2.1_territorial == 0),
                       control = glmerControl(optimizer = "bobyqa",
                                              optCtrl = list(maxfun = 2e5))) #filter by the non-territorial campaigns
#WARNING: A convergence issue occured. Reestiamte a model with a warm start
all_fits_model2_noterr <- allFit(model2_noterr)
ok_fits  <- all_fits_model2_noterr[summary(all_fits_model2_noterr)$which.OK]
llik_vals <- sapply(ok_fits, logLik)
model2_noterr  <- ok_fits[[which.max(llik_vals)]]
#model2_noterr@optinfo$conv$lme4 #model converged

ova_egip_africa_m2  <- sim_ova_range(model2_africa,        
                                     focal_var = "epr_highest_rank", 
                                     nsim = simulation_number,
                                     vcov_matrix = vcov(model1_africa))$result %>% mutate(type = "MENA & Sub-Saharan Africa")
ova_egip_noterr_m2  <- sim_ova_range(model2_noterr,        
                                     focal_var = "epr_highest_rank", 
                                     nsim = simulation_number,
                                     vcov_matrix = vcov(model1_noterr))$result %>% mutate(type = "Excl. territorial campaigns")

ova_model2_robustness <- bind_rows(
  ova_egip_africa_m2,
  ova_egip_noterr_m2,
) %>%
  mutate(type = factor(type, levels = c(
    "MENA & Sub-Saharan Africa",
    "Excl. territorial campaigns"
  )))


ggplot(ova_model2_robustness,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df$NVC2.1_VIOL) / nrow(df), 2),  # a benchmark probability dashed line
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data = df %>% drop_na(epr_IFI_l),
    aes(x = epr_IFI_l),
    inherit.aes = FALSE,
    color = "#C0392B", alpha = 0.3,
    sides = "b"          # bottom only
  ) + 
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )

```
