---
title: "**Guns or Gandhi: Ethnic Elite Composition and Dissident Tactics in Civil Conflicts**"
author: 
  - "Ruslan Guseinov"
  - "Maaike Arienne Zeelen"
output:
    bookdown::pdf_document2:
 #   toc: true
    number_sections: false
  #pdf_document:
  #  toc: true
linestretch: 2
indent: true
fontsize: 12pt
geometry: left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm
date: "`r Sys.Date()`"
bibliography: references.bib
header-includes:
  - \setcounter{secnumdepth}{0}
  - \usepackage{amsthm}
  - \newtheorem{hyp}{Hypothesis}
abstract: |
  This paper is very cool blbablalblab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,    # hide code from output
  warning = TRUE,    # suppress warnings
  message = FALSE,     # suppress package loading messages
  include = FALSE # hide code
)

p_needed <- c(
  "tidyverse",        # dplyr, ggplot2, tidyr throughout
  "lme4",             # glmer for RE logit
  "MASS",             # mvrnorm in OVA simulation functions
  "kableExtra",       # kbl, kable_styling for all tables
  "patchwork",        # combining histogram + density plot
  "bookdown",         # pdf_document2, cross-referencing
  "countrycode",      # datacollection.R country coding
  "pROC",             # AUC-ROC model fit
  'modelsummary'      # For summary tables
)
packages <- rownames(installed.packages())
p_to_install <- p_needed[!(p_needed %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install) 
}
sapply(p_needed, require, character.only = TRUE)
options(warn = -1)

#download.file(
#  "https://www.zotero.org/styles/apa",
#  destfile = "apa.csl"
#) #download APA citation style

stargazer_opt <- ifelse(knitr::is_latex_output(), "latex", "html")
```

```{r collect-data}
# USE THESE FUNCTIONS FOR DATA COLLECTION
# DATASETS ARE ALREADY STORED IN THE FOLDER
#suppressWarnings({source("datacollectionscripts/EPR_construction.R")}) #Tibet (code 711) is dropped
#suppressWarnings({source('datacollectionscripts/datacollection.R')}) 
```

```{r functions}
source('functions/simulation_functions.R') #use pre-defined simulation functions
source('functions/ova_sim_functions.R') #use observed value approach (OVA) simulation functions

simulation_number = 1000
```

```{r load-data}
lost_campaigns_table <- read.csv("datasources/lost_campaigns.csv") %>%
  dplyr::select(-1)
df_final_regr <- readRDS('datasources/df_essay_regr.rds') # read a dataset
# RESTRICT THE DATASET TO ONLY ONSETS
# TEST THE HYPOTHESES ABOUT THE CHOICE OF METHODS
df_final_methods <- df_final_regr %>% 
  filter(onset_type %in% c('Nonviolent', 'Violent'))

# Add variables for a Within-Between Random Effect modelling
df_final_methods <- df_final_methods %>%
  group_by(cow) %>%
  mutate(
    mean_IFI     = mean(epr_IFI_l, na.rm = TRUE), 
    mean_EFI     = mean(epr_EFI_l, na.rm = TRUE),
    demeaned_IFI = epr_IFI_l - mean_IFI,
    demeaned_EFI = epr_EFI_l - mean_EFI
  ) %>%
  ungroup()
```


\section{Introduction}

\section{Violent and Non-violent Campaign Choice}

Previous literature has established that civil resistance campaigns and violent campaigns are often a strategic choice made before the campaign [@schock2005unarmed; @Cunninghametal2017]. One strand of this literature emphasizes the role of political opportunities in shaping these choices (). Campaign leaders evaluate the political environment and assess the likely consequences of alternative tactics, including the prospects for mobilization, repression, elite defections, and policy concessions. The underlying assumption is that dissident organizations are more likely to adopt the tactic they expect to offer the greatest probability of achieving their political objectives. The following paragraphs summarize key political opportunities that shape the relative attractiveness of nonviolent and violent resistance.

A nonviolent campaign is a large-scale collective effort that uses nonviolent methods to achieve major political goal. Such goals may include overthrowing an incumbent government or achieving territorial secession. Common tactics include strikes, boycotts, and sit-ins.

Nonviolent campaigns rely on broad participation and cooperation from large segments of society. Their effectiveness therefore depends on their ability to mobilize supporters, disrupt normal political activity and weaken the regime's sources of support.

As nonviolent resistance depends on mass participation, repression can substantially reduce its effectiveness. Arrests, violence, and intimidation increase the costs of participation and may discourage potential supporters from joining collective action (). Repression can also disrupt organizational networks by targeting leaders and activists who coordinate protest activities ().

[Paper] showed that a key mechanism through which nonviolent campaigns succeed is by inducing defections among regime supporters. When political elites, bureaucrats, or members of the security forces withdraw their support from the incumbent government, the regime's capacity to govern and repress dissent is weakened ().

In addition to limiting repression, regimes may choose to negotiate or make concessions in response to nonviolent mobilization. When dissidents expect that peaceful collective action can generate meaningful concessions, the expected benefits of nonviolent tactics increase relative to violent alternatives.

Nonviolent resistance is therefore most attractive when dissidents expect high levels of participation, a substantial likelihood of elite or security-force defections, limited repression, and a reasonable prospect of political concessions.


Violent campaigns involve the use of armed force or coercive violence to achieve large political objectives. Unlike nonviolent campaigns, violent resistance does not depend on broad participation in the same way and can be sustained by smaller, more centralized organizations.

Because violent campaigns do not rely on mass mobilization, they may become relatively more attractive when conditions make large-scale participation in nonviolent resistance difficult. In particular, severe repression can raise the costs of open collective action, reducing the feasibility of sustained nonviolent mobilization.

Violent resistance may also be more likely when opportunities for elite or security-force defection are limited. When regimes maintain cohesive ruling coalitions and reliable coercive institutions, nonviolent campaigns are less able to generate the defections that often undermine incumbent governments.

Consequently, political environments characterized by strong elite cohesion, effective repression, and low prospects for defection tend to increase the relative attractiveness of violent forms of resistance.


\section{Ethnic Composition and Campaign Tactic}

Political opportunities play a central role in shaping whether dissident campaigns adopt violent or nonviolent strategies. Campaigns evaluate their prospects of success based on expectations about elite defections, repression, and political concessions, and choose the tactic that appears most likely to achieve their objectives.

This study argues that variation in these political opportunities is structured by the ethnic composition of ruling elites. Ethnic fragmentation shapes elite cohesion and thereby influences the likelihood of defection, the effectiveness of repression, and the probability of concessions.

Ethnically heterogeneous ruling coalitions are more vulnerable to elite defection because political loyalty is often embedded in ethnic networks that provide alternative bases of support outside the regime. This increases the likelihood that elites withdraw support under conditions of sustained mobilization.

Ethnic divisions can also weaken coercive capacity by generating coordination problems within security forces and governing institutions, reducing the state’s ability to carry out sustained repression.

In addition, fragmented ruling coalitions may be more likely to pursue negotiated settlements, as internal divisions increase the risks associated with prolonged conflict and make concessions a more attractive option for regime survival.

Taken together, ethnic fragmentation increases the likelihood of elite defections, weakens repression, and increases the probability of concessions. These conditions make nonviolent resistance more effective relative to violent forms of contention.

Accordingly, we expect that dissident campaigns facing more ethnically fragmented ruling coalitions will be more likely to adopt nonviolent tactics.

<!-- Hypotheses are automatically numerated -->
\begin{hyp}
Greater ethnic fragmentation in ruling elites increases the likelihood that dissident campaigns adopt nonviolent tactics rather than violent tactics.
\end{hyp}

Variation in political opportunities may also arise from differences in the distribution of power among ethnic groups within ruling coalitions. More equally powerful groups possess greater autonomy and capacity to defect, making elite cohesion weaker than in systems dominated by a single group.

Greater equality in the distribution of power among ethnic groups therefore increases the likelihood of elite defection and reduces the effectiveness of repression.

From this mechanism, I derive a second hypothesis:

\begin{hyp}
Greater equality in the distribution of power among ethnic groups in the ruling elite increases the likelihood that dissident campaigns adopt nonviolent tactics.
\end{hyp}


\section{Data and Methods}
\subsection{Data description}
Our study employs a panel dataset of `r nrow(df_final_methods)` so-called "maximalist campaigns", a concept first introduced by Chenoweth and Stephan in *Why Civil Resistance Works* book [@chenoweth2011civil]. Following their definition, a campaign is "a series of observable, continuous, purposive mass tactics or events in pursuit of a political objective’, with maximalist objectives encompassing regime change, self-determination, secession, and the expulsion of foreign occupations [@DVN/MHOXDV_2019]. We rely on Nonviolent and Violent Campaigns and Outcomes (NAVCO) 2.1 dataset, which records 384 maximalist campaigns from 1945 to 2013^[The most significant reason why the number of campaigns we have in the panel dataset differs from the total number of maximalist campaigns concerns country coding. We construct a panel dataset of indepedent country-years based on Correlates of War (CoW) list of independent states. We merge our panel data using NAVCO-assigned CoW country codes. However, in some cases, countries of a campaign location are not independent at the campaign onset year and thus lost from the data we use in the analysis. For example, there is a campaign geographically located in Algeria in 1952, though Algeria gained independence only in 1962. A naive solution would be assigning socio-economic characteristics of the "target" country - France in case of Algeria, though it would severely bias the results as if the campaign took place in France. We partly address this issue by excluding territorial campaigns in the robustness check section. Table \@ref(tab:lost-campaigns-appendix) in Appendix C lists all these lost campaigns.].

The central conceptual distinction is between violent and nonviolent repertoires of contention. Following NAVCO coding rules, nonviolent tactics ‘do not directly threaten or harm the physical wellbeing of the regime, its agents, or its citizens’ and the violent repertoire ‘involves the use of force to physically harm or threaten to harm the opponents’ [@DVN/MHOXDV_2019]. The categories are mutually exclusive. A prominent example of a nonviolent campaign is the German 1989 *Die Wende* and the Cuban Revolution exemplifies a violent campaign. The violence/nonviolence distinction is well-established in the peace science literature and some significant findings have been made using NAVCO dataset [@celestino2013fresh; @manekin2022effective; @cunningham2013understanding]. Violence is operationally coded as a binary variable (1 = a campaign takes a violent form, 0 = a campaign takes a nonviolent form).

Our key independent variables are drawn from the Ethnic Power Relations (EPR) Core Dataset [@vogt2015integrating]. EPR provides information on ethnic groups and their relative access to power. Following Hendrix and Salehyan [-@hendrix2019ethnicity], we operationalize the ethnic ruling elite fragmentation via the *Included fractionalization index (IFI)* and the *Excluded fractionalization index (EFI)*. IFI is "a Herfindahl-like index of all groups counted as ‘included’ (i.e. having access to executive power) during a given country-year", whilst EFI is "a Herfindahl-like index of all groups counted as ‘excluded’ during a given country-year". Thus, a higher IFI indicates a more concentrated ethnic ruling elite and a lower IFI shows that ethnic power is distributed more equally, across many groups. Instead of relying on raw counts, IFI and EFI are more theoretically motivated. If a raw count treats two countries with the ethnic groups in power count of, say, 3 equally, IFI and EFI account for ethnic power distributions.

Another important source of data is the Ethnic Groups in Contention (EGC), a dataset first introduced by Thurber in 2018 [@thurber2018ethnic]. It links NAVCO to EPR and provides additional information on ethnic groups that participated in maximalist campaigns. In a 2026 article, the EGC dataset was updated for NAVCO 2.1, allowing us to use the newest version for our research question [@Bivens26]. What we are interested in is whether any ethnic group in power in a specific country participated in the maximalist campaign. Obviously, ethnic groups that are systematically discriminated could mobilize against the ruling regime given their ethnic (and possibly other) grievances, though empowered ethnic groups could also participate in the campaign, supporting the excluded masses - a potential source of significant confounding. Even though such scenarios sound unrealistic, there are `r sum(df_final_methods$egip_participated, na.rm=TRUE)` where at least one ethnic group in power participated in the campaign. However, in only `r sum(df_final_methods$egip_participated_claimed, na.rm=TRUE)` did such participating ethnic groups in power made an ethnic claim, displaying rather an opportunistic behavior in anticipation of the maximalist campaign success than a sincere will to express ethnic grievances.

<!-- Inspired by Bivens and Thurber [-@Bivens26], we focus on the Ethnic Groups in Power (EGIP) and Marginalized Ethnic Groups (MEG^[We also refer to MEG as "excluded groups", these concepts are used synonymously]) counts. EGIP encompasses groups that exercise political power either via maintaining ethnic dominance or through power-sharing arrangements with junior partners, whilst MEG covers groups that are politically excluded from decision making, including both powerless and purposely discriminated groups. Due to theoretical expectations that a marginal effect of the ethnic groups count on the tactic choice diminishes at higher values, we apply the natural logarithm transformations to both EGIP and MEG counts in the empirical analysis. -->

To address endogeneity and block non-causal paths, we control for a standard set of confounders: GDP per capita in constant PPP dollars [@gapminderGDPCapitaConstant2024]; logged population size [@unitednationsWorldPopulationProspects2024]; executive corruption scores and polyarchy (electoral democracy) scores [@coppedgeVDemCodebookV142024]; regime duration in years [@coppedgeVDemCodebookV142024]; peace years (years elapsed since the last maximalist campaign in each country); EFI to account for a possible confounding of excluded ethnic groups as well; we also control whether at least one ethnic group in power participated in the campaign with a binary variable. To address reversed causality, all the independent variables are one-year lagged^[Even though we find it unconvincing that a choice of one tactic over the other could affect IFI and EFI].

\subsection*{Descriptive statistics}
```{r descriptive-stats, echo=FALSE, include = TRUE}
summarydata <- df_final_methods %>%
  mutate(
    Tactics = case_when(
      onset_type == "Nonviolent" ~ "Nonviolent",
      TRUE                       ~ "Violent"
    )
  ) %>%
  dplyr::select(
    Tactics,
    "Included Fractionalization Index"                = epr_IFI_l,
    "Excluded Fractionalization Index"                = epr_EFI_l,
    "Ethnic Group in Power Participation" = egip_participated_l,
    "ln(GDP per capita)" = log_gdp_pcap_l,
    "ln(Population)"     = pop_log_l,
    "Polyarchy"          = v2x_polyarchy_l,
    "Executive Corruption" = v2x_execorr_l,
    "ln(Regime Duration)"  = log_v2regdur_l,
    "Peace Years"          = peace_years_l
  )

datasummary_balance(
  ~Tactics,
  data    = summarydata,
  title   = "Ethnic and political attributes of NAVCO 2.1 campaign onsets",
  label   = "descriptive-stats",
  output  = "kableExtra",
  fmt     = 3,          # 3 decimal places
  dinm    = FALSE       # no difference-in-means column
) %>%
  kable_styling(
    latex_options = c("scale_down"),
    full_width    = FALSE
  ) %>%
  footnote(
    general           = "All variables one-year lagged.",
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable    = TRUE
  ) %>% kableExtra::kable_styling(latex_options = c("hold_position"))
```


Table \@ref(tab:descriptive-stats) presents descriptive statistics for all variables included in the analysis between nonviolent and violent campaigns. The table shows a balance between violence and nonviolence instances. Nonviolent campaigns have a slightly higher mean Included Fractionalization Index value, while violent campaigns face a slightly higher mean Excluded Fractionalization Index value. For control variables, such as GDP and polyarchy, the table shows some degree of variation between violent and nonviolent campaigns. In general, data missingness is not a significant issue, though there is missing data on ethnic groups for all campaigns in countries where data collection is complicated and ethnic lines are profound, such as Yemen and Somalia^[In general, only `r round(sum(is.na(df_final_methods$epr_IFI_l)) / nrow(df_final_methods) * 100, 2)`% of IFI values and `r round(sum(is.na(df_final_methods$epr_EFI_l)) / nrow(df_final_methods) * 100, 2)`% of EFI values are missing]. Figure \@ref(fig:epr-ifi) shows the distribution of the included fractionalization index with respect to the tactic choice, visualized by both a histogram and a density plot. At low IFI values, violent tactics are more frequent and with the growing IFI the nonviolent tactic is prevalent. However, with the growing IFI both tactic choices occur less frequently, reflecting the mass anti-regime mobilization logic - with the growing elite cohesion, political opportunities of rebellion are decreasing, regardless of the repertoire of contention. This visualization shows some degree of variation of 1) the Included Fractionaliztion Index 2) between violent and nonviolent choice.

```{r epr-ifi, fig.cap="Included fractionalization index by Repertoire Choice: a histogram and a density plot", fig.align = 'center', out.width="85%", include = TRUE}
# Histogram
p_hist <- df_final_methods %>%
  ggplot(aes(x = epr_IFI_l, fill = onset_type)) +
  geom_histogram(binwidth = 0.1, position = "dodge",
                 color = "white", linewidth = 0.3) +
  scale_fill_manual(values = c("Nonviolent" = "#2C86C7",
                               "Violent"    = "#C0392B")) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(x = "IFI (0 = fully fragmented, 1 = fully concentrated)",
       y = "Count",
       fill = NULL) +
  theme_bw() +
  theme(legend.position  = "none",
        panel.grid.minor = element_blank())

# Density plot
p_dens <- df_final_methods %>%
  ggplot(aes(x = epr_IFI_l, fill = onset_type, color = onset_type)) +
  geom_density(alpha = 0.25, linewidth = 0.8, bw = 0.06) +
  scale_fill_manual(values  = c("Nonviolent" = "#2C86C7",
                                "Violent"    = "#C0392B"), guide   = "none") +
  scale_color_manual(values = c("Nonviolent" = "#2C86C7",
                                "Violent"    = "#C0392B"), guide   = "none") +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2)) +
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)",
       y     = "Density",
       fill  = NULL,
       color = NULL) +
  theme_bw() +
  theme(legend.position  = "bottom",
        panel.grid.minor = element_blank())

# Combine with patchwork
(p_hist + p_dens) + 
  plot_layout(axis_titles = "collect", guides = 'collect') &
  theme(
    legend.position        = "bottom",
    legend.justification   = "center",
    legend.box.just        = "center",
    panel.grid.minor       = element_blank()
  )
```

\subsection*{Modeling approach}

```{r within-sd-re}
# calculate a standard deviation of IFI and EFI grouped by countries
# to justify RE over FE
within_sd <- df_final_methods %>%
  group_by(cow) %>%
  summarise(
    sd_IFI = sd(epr_IFI_l, na.rm = TRUE),
    sd_EFI = sd(epr_EFI_l, na.rm = TRUE),
    n_obs  = n()
  )
```


Given a binary dependent variable and a balanced sample -`r nrow(df_final_methods[df_final_methods$NVC2.1_VIOL == 1,])` of violent and `r nrow(df_final_methods[df_final_methods$NVC2.1_VIOL == 0,])` of nonviolent onsets - we estimate a logit model. As we do not predict the duration of events, there is no need for advanced binary choice models such as a complementary log-log. To account for a nested structure of our data, where campaigns are nested within countries, we add a country-level random varying intercept in all models. 

Including a country fixed effect to control for the time-invariant unobserved factors such as culture or geography is theoretically appealing and common in the peace science literature [@butcher2016manufacturing; @ives2022greed; @fearon2003ethnicity; @jung2025motivating]. However, given that ethnic structures are almost constant *within* countries and vary predominantly *between* countries - the variation that a country fixed effect absorbs, adding it is methodologically inappropiate. Exploratory data analysis showed that the median within-country standard deviation of both IFI and EFI rounds to `r round(median(within_sd$sd_IFI, na.rm=TRUE), 2)` (up to two decimal points without rounding), meaning that the FE identification would test on a very small sample of countries that experienced ethnic configuration changes between maximalist campaigns. Random effects exploit both within- and between-group variations, which is appropriate for our research question and data structure. Moreover, a random-effects estimator introduces "a compromise between the fixed-effects and pooled models", though we acknowledge that the coefficients can be partially biased in case varying intercepts correlate with the regressors [@clark2015should]. Despite this potential bias, RE is still a more preferable choice over FE.

For Hypothesis 1, our model specification is:
$$
\begin{aligned}
& \text{violent}_{it} \sim \text{Bernoulli}(\pi_{it}) \\[0.3em]
& \log\left(\frac{\pi_{it}}{1-\pi_{it}}\right) = \beta_0 
+ \beta_1 (\text{IFI})_{it-1} 
+ \beta_2 (\text{EFI})_{it-1} 
+ \mathbf{X}'_{it-1}\boldsymbol{\gamma} + \alpha_{[j]i} \\[0.3em]
& \alpha_j \sim \mathcal{N}(\mu_\alpha, \sigma^2_\alpha)
\end{aligned}
$$
where $\mathbf{X}_{it-1}$ is the vector of covariates and $\alpha_{[j]i}$ is a country varying intercept.

For uncertainty calculations, we follow a King's simulation approach [@king2000making]. This approach is also well-established in peace science literature [@butcher2016manufacturing]. Instead of reporting coefficients and their significance, we calculate predicted probabilities as the main quantity of interest via an observed value approach (OVA), meaning we simulate the predicted probabilities for all observations and then calculate the mean of them^[We report coefficients only in the cross-validation robustness check]. We simulate over the full available range of the vocal variable values. For the essay, we define `sim_ova_range` function for the OVA simulations, allowing to calculate predicted probabilities in just four lines of code. It also allows for different CI levels, seeds, number of values and focal variable ranges at the researcher's discretion.

\subsection{Results}

```{r h1-modeling}
# estimate a null model
null_model <- glmer(NVC2.1_VIOL ~ 
                      1 + 
                      (1 | cow), # random intercept by country
                    data   = df_final_methods,
                    family = binomial("logit"))

#define a formula for h1
formula_h1 <- NVC2.1_VIOL ~ epr_IFI_l + epr_EFI_l + egip_participated_l +
                    log_gdp_pcap_l + pop_log_l + v2x_polyarchy_l + v2x_execorr_l +
                    log_v2regdur_l + peace_years_l + (1 | cow)  # random intercept by country

model1 <- glmer(formula_h1,
                  data = df_final_methods,
                  family = binomial(link = "logit"),
                  control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) #apply an optimizer for convergence
```

```{r h1-qoisim}
ova_egip_full_m1  <- sim_ova_range(model1,        
                                 focal_var = "epr_IFI_l", 
                                 nsim = simulation_number,
                                 vcov_matrix = vcov(model1))$result %>% mutate(type = "Full Sample")
```

```{r h1-plot, fig.cap="Simulated predicted probabilities of a violent tactic choice over a range of included fractionalization index values", fig.align = 'center', out.width="85%", include = TRUE}
ggplot(ova_egip_full_m1,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2), # a benchmark probability dashed line
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data = df_final_methods %>% drop_na(epr_IFI_l),
    aes(x = epr_IFI_l),
    inherit.aes = FALSE,
    color = "#C0392B", alpha = 0.3,
    sides = "b"          # bottom only
  ) + 
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```
To test the first hypothesis, we estimate a logit model of the violent tactic choice, according to the H1 model specification. Figure \@ref(fig:h1-plot) shows the simulated predicted probabilities of a violent tactic choice over a full available range of observed IFI values. Despite a downward trend - from `r round(ova_egip_full_m1$mean[1], 2)` at IFI equal to `r round(ova_egip_full_m1$focal_value[1], 2)` to `r round(ova_egip_full_m1$mean[50], 2)`  at IFI equal to `r round(ova_egip_full_m1$focal_value[50], 2)` - confidence intervals that encompass 0.5 do not allow us to claim that with the increasing IFI the rebels would prefer violence over nonviolence. Instead of zero probability, our benchmark probability is approx. `r round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2)` - the sample proportion of violent campaigns in the whole sample.

```{r h1-train-test sampling, include = TRUE}
set.seed(666)  # same seed as simulations for reproducibility

n_total <- nrow(df_final_methods) #define the number of observations
train_idx <- sample(1:n_total, size = floor(2/3 * n_total), replace = FALSE) #slice two thirds into the train set
test_idx  <- setdiff(1:n_total, train_idx) #all other observations into the test

df_final_methods_train <- df_final_methods[train_idx, ] # train dataset

model1_train  <- glmer(formula_h1,        # random intercept by country
                  data = df_final_methods_train,
                  family = binomial(link = "logit"),
                  control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5)))

trained_cows <- rownames(ranef(model1_train)$cow) # what countries are estimated?
df_final_methods_test <- df_final_methods[test_idx, ] %>%
  filter(as.character(cow) %in% trained_cows) %>%  #test is only with the trained countries
  drop_na(epr_IFI_l, epr_EFI_l, egip_participated_l, log_gdp_pcap_l, pop_log_l, v2x_execorr_l, v2x_polyarchy_l,
          log_v2regdur_l, peace_years_l)

pred_test_prob <- predict(model1_train, newdata = df_final_methods_test, type = "response")  #predict on the test dataset
actual_test <- ifelse(df_final_methods_test$NVC2.1_VIOL == 1,
                      "Violent", "Nonviolent")
pred_test   <- ifelse(pred_test_prob >= 0.5, "Violent", "Nonviolent") # 0.5 as the prediction threshold

roc_obj  <- roc(response  = ifelse(actual_test == "Violent", 1, 0),
                predictor = pred_test_prob,
                quiet     = TRUE)
auc_val  <- as.numeric(auc(roc_obj))

tp <- sum(pred_test == "Violent"    & actual_test == "Violent")
tn <- sum(pred_test == "Nonviolent" & actual_test == "Nonviolent")
fp <- sum(pred_test == "Violent"    & actual_test == "Nonviolent")
fn <- sum(pred_test == "Nonviolent" & actual_test == "Violent")

# calculate metrics
accuracy  <- (tp + tn) / (tp + tn + fp + fn)
precision <- tp / (tp + fp)
recall    <- tp / (tp + fn)   
f1        <- 2 * (precision * recall) / (precision + recall)

conf_df <- data.frame(
  ` `            = c("Predicted Nonviolent", "Predicted Violent"),
  `Actual Nonviolent` = c(as.character(tn), as.character(fp)),
  `Actual Violent`    = c(as.character(fn), as.character(tp)),
  check.names = FALSE
)

metrics_df <- data.frame(
  ` `                 = c("Accuracy", "Precision", 
                           "Recall",  "F1 Score", "AUC"),
  `Actual Nonviolent` = c(sprintf("%.3f", accuracy), 
                           sprintf("%.3f", precision),
                           sprintf("%.3f", recall), 
                           sprintf("%.3f", f1),
                           sprintf("%.3f", auc_val)),
  `Actual Violent`    = c("", "", "", "", ""),
  check.names = FALSE
)

bind_rows(conf_df, metrics_df) %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Confusion Matrix and Predictive Performance — Model 1",
    label    = "conf-matrix-h1",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(2, extra_css = "border-bottom: 2px solid black;") %>%  # separator line
  pack_rows("Confusion Matrix", 1, 2) %>%
  pack_rows("Performance Metrics", 3, 7)
```

Despite an insignificant effect of IFI on a violent choice, the model as a whole demonstrates a strong predictive performance. We divided the dataset into a train and test sample, estimated a logit model on the train sample and predicted a tactic choice on the test set. Table \@ref(tab:conf-matrix-h1) reports the confusion matrix and performance metrics. With an accuracy of `r round(accuracy, 2)` and F1 score of `r round(f1, 2)`, our model in general predicts tactic choice well above a coin toss, predicting violence a bit better than the nonviolent choice. Another measure of model fit is the area under the ROC curve (AUC), showing whether our model predictions of a violent choice are better than a random draw. The AUC score of `r round(auc_val, 2)` indicates a good model fit. Figure \@ref(fig:h1-auc-roc-appendix) plots an AUC-ROC curve for this model. An important implication here is that the model as a whole (including the specification itself, a set of covariates and a country varying intercept) captures meaningful variation in the outcome, even though the IFI does not significantly predict a violent choice. Such results reflect a real absence of the IFI effect rather than a model misspecification.

\subsection{Robustness checks}
We conduct four robustness checks. First, given that civil conflicts and a tactic choice are regionally clustered, we estimate an additional model on the Middle East and North Africa (MENA) and Sub-Saharan Africa countries subsample, where violent campaigns are more expected than in other regions. For the full sample, there are `r round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods) * 100, 1)`% of violent campaigns, whilst for MENA \& Sub-Saharan Africa there are `r round(sum(df_final_methods[df_final_methods$region %in% c('Middle East & North Africa', 'Sub-Saharan Africa'),]$NVC2.1_VIOL) / nrow(df_final_methods[df_final_methods$region %in% c('Middle East & North Africa', 'Sub-Saharan Africa'),]) * 100, 1)`% of them. Second, territorial campaigns could be more ethnicity-driven and secessionist civil wars along ethnic lines are widely researched in the academic literature [@collier2006political; @denny2014ethnicity; @toft2012self]. For this reason, we estimated three separate models 1) on the MENA & Sub-Saharan Africa subsample, 2) excluding territorial campaigns and 3) both on the MENA & Sub-Saharan Africa sub-sample with excluded territorial campaigns. Finally, following Bell and Jones [-@bell2015explaining], we estimate an alternative Within-Between Random Effect (REWB) model specification. REWB presents a main independent variable in two forms: its demeaned form and a group mean. De-meaning allows to estimate a *within* (essentially fixed) effect and save the underlying assumption of the varying intercept model. An important note here is that we do not compare the model fit of these specifications because of the different samples, which make the models incomparable in terms of model fit measures.

```{r h1-robustness-modelling}
model1_africa <- glmer(formula_h1,
    family = binomial(link = "logit"),
    data = df_final_methods %>% filter(region %in% c('Middle East & North Africa', 'Sub-Saharan Africa')),
                  control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) # filter by the regionss

model1_noterr <- glmer(formula_h1,
    family = binomial(link = "logit"),
    data = df_final_methods %>% filter(NVC2.1_territorial == 0),
                  control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5))) #filter by the non-territorial campaigns

model1_africa_noterr <- glmer(formula_h1,
    family = binomial(link = "logit"),
    data = df_final_methods %>% filter(region %in% c('Middle East & North Africa', 'Sub-Saharan Africa')) %>% filter(NVC2.1_territorial == 0),
                  control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5)))
```

```{r h1-robustness-qoisim}
ova_egip_africa_m1  <- sim_ova_range(model1_africa,        
                                 focal_var = "epr_IFI_l", 
                                 nsim = simulation_number,
                                 vcov_matrix = vcov(model1_africa))$result %>% mutate(type = "MENA & Sub-Saharan Africa")
ova_egip_noterr_m1  <- sim_ova_range(model1_noterr,        
                                 focal_var = "epr_IFI_l", 
                                 nsim = simulation_number,
                                 vcov_matrix = vcov(model1_noterr))$result %>% mutate(type = "Excl. territorial campaigns")
ova_egip_africa_noterr_m1  <- sim_ova_range(model1_africa_noterr,        
                                 focal_var = "epr_IFI_l", 
                                 nsim = simulation_number,
                                 vcov_matrix = vcov(model1_africa_noterr))$result %>% mutate(type = "MENA & Sub-Saharan Africa & Excl. terr.")

ova_model1_robustness <- bind_rows(
  ova_egip_full_m1,
  ova_egip_africa_m1,
  ova_egip_noterr_m1,
  ova_egip_africa_noterr_m1
) %>%
  mutate(type = factor(type, levels = c(
    "Full Sample",
    "MENA & Sub-Saharan Africa",
    "Excl. territorial campaigns",
    "MENA & Sub-Saharan Africa & Excl. terr."
  )))
```

```{r h1-robustness-plot, fig.cap="Simulated predicted probabilities of a violent tactic choice over a range of IFI values - robustness checks results", fig.align = 'center', out.width="85%", include = TRUE}
ggplot(ova_model1_robustness,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2),  # a benchmark probability dashed line
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data = df_final_methods %>% drop_na(epr_IFI_l),
    aes(x = epr_IFI_l),
    inherit.aes = FALSE,
    color = "#C0392B", alpha = 0.3,
    sides = "b"          # bottom only
  ) + 
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```

We calculate simulated predicted probabilities over the full available range of IFI values for all three robustness check models and summarize the findings on one plot. Figure \@ref(fig:h1-robustness-plot) shows the results along three robustness specifications, including the full-sample baseline. The only partially significant finding is in the model which excludes territorial campaigns (this sub-sample mostly includes anti-regime campaigns). For low IFI values, campaigns that exclude territorial goals/claims have a higher chance of being nonviolent and the rebels slowly push towards a higher likelihood of violence with the growing IFI value. This (though cautious) finding can be explained by the logic of mass anti-regime mobilization, where a more concentrated ethnic elite decreases the benefits of nonviolent resistance and pushes the rebels toward violent repertoires of contention. For the MENA \& Sub-Saharan Africa subsample (both with and without territorial campaigns) there is no systematic difference between violence and nonviolence across the growing IFI value. Figure \@ref(fig:h1-robustness-plot-appendix) plots simulated predicted probabilities for the baseline and the REWB model, demonstrating practically identical results (see Appendix A). In general, our null result is robust^[We do not calculate performance metrics for the robustness check specifications due to identification issues and low sample sizes, caused by simultaneous regional/campaign type and train-test subsampling.].


```{r cross-validation-function-glm}
cross_validation_fun_glm <- function(validation_number = 10,
                                   data,
                                   dep_var) {
  n <- validation_number # set the number of validations
  seed_vectors <- seq_len(n) # a set of seeds (increasing number of validations)

  # model specification
  formula_obj <- as.formula(
    paste(dep_var, "~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l +
    log_epr_egippop_l + log_gdp_pcap_l + pop_log_l + youthbulge_l + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l + log_v2regdur_l + peace_years_l + egip_participated + egip_claimed + as.factor(cow)")
  )

  # balance the sample by realizations and non-realizations
  data_realizations    <- data[data[[dep_var]] == 1, ]
  data_nonrealizations <- data[data[[dep_var]] == 0, ]

  # Get substantive coefficient names — drop country FE
  test_fit   <- glm(formula_obj, family = binomial("logit"), data = data)
  all_names  <- names(coef(test_fit))
  keep_idx   <- !grepl("as.factor\\(cow\\)", all_names)  # exclude FE
  coef_names <- all_names[keep_idx]
  n_coefs    <- length(coef_names)

  coefs_matrix <- matrix(NA, nrow = n, ncol = n_coefs)
  var_matrix   <- matrix(NA, nrow = n, ncol = n_coefs)
  colnames(coefs_matrix) <- coef_names
  colnames(var_matrix)   <- coef_names

  # loop through validations
  for (i in 1:n) {
    set.seed(seed_vectors[i]) # set a seed

    samp1 <- sample(1:nrow(data_realizations),
      size = floor(2 / 3 * nrow(data_realizations))
    ) # sample realizations
    samp0 <- sample(1:nrow(data_nonrealizations),
      size = floor(2 / 3 * nrow(data_nonrealizations))
    ) # sample non-realizations

    sample_data <- rbind(
      data_realizations[samp1, ],
      data_nonrealizations[samp0, ]
    ) # combine the data

    model  <- glm(formula_obj, family = binomial("logit"), data = sample_data)
    
    vec_coefs <- coef(model)
    vec_var   <- diag(vcov(model))

    # Keep only substantive coefficients — drop FE
    keep      <- !grepl("as.factor\\(cow\\)", names(vec_coefs))
    vec_coefs <- vec_coefs[keep]
    vec_var   <- vec_var[keep]

    # Align by name
    idx <- match(names(vec_coefs), coef_names)
    idx_valid <- !is.na(idx)
    coefs_matrix[i, idx[idx_valid]] <- vec_coefs[idx_valid]
    var_matrix[i,   idx[idx_valid]] <- vec_var[idx_valid]
  }

  # Apply the King (2001) rule

  # combined coefficients
  mean_coef <- colMeans(coefs_matrix, na.rm = TRUE)
  # within-sample variance
  W <- colMeans(var_matrix, na.rm = TRUE)
  # between-sample variance
  B <- apply(coefs_matrix, 2, var, na.rm = TRUE)
  # total variance
  T_var <- W + (1 + 1 / n) * B
  # combined standard errors
  se_combined <- sqrt(T_var)

  # inference
  z_value <- mean_coef / se_combined
  p_value <- 2 * (1 - pnorm(abs(z_value)))

  stars <- ifelse(p_value < 0.001, "***",
    ifelse(p_value < 0.01, "**",
      ifelse(p_value < 0.05, "*",
        ifelse(p_value < 0.1, ".", "")
      )
    )
  )

  results <- data.frame(
    Variable = coef_names,
    Coefficient = round(mean_coef, 6),
    `Std. Error` = round(se_combined, 6),
    `z value` = round(z_value, 3),
    `Pr(>|z|)` = round(p_value, 4),
    Sig = stars,
    check.names = FALSE
  )

  # return a list with the final results, coefficients and ses separately
  return(list(
    results = results,
    combined_coefs = mean_coef,
    combined_se = se_combined,
    coefs_matrix = coefs_matrix,
    var_matrix = var_matrix
  ))
}
```

```{r cross-validation-function-glmer}
cross_validation_fun_glmer <- function(validation_number = 50,
                                 data,
                                 dep_var) {
  
  formula_obj <- as.formula(
    paste(dep_var, "~ epr_IFI_l + epr_EFI_l + egip_participated_l +
                    log_gdp_pcap_l + pop_log_l + v2x_polyarchy_l + v2x_execorr_l +
                    log_v2regdur_l + peace_years_l  +  
                    (1 | cow)")
  )
  
  data_realizations    <- data[data[[dep_var]] == 1, ]
  data_nonrealizations <- data[data[[dep_var]] == 0, ]
  
  test_fit   <- glmer(formula_obj, family = binomial("logit"), data = data,
                      control = glmerControl(optimizer = "bobyqa",
                                            optCtrl   = list(maxfun = 2e5)))
  coef_names <- names(fixef(test_fit))
  n_coefs    <- length(coef_names)
  
  # pre-allocate for target number of CONVERGED models
  coefs_matrix <- matrix(NA, nrow = validation_number, ncol = n_coefs)
  var_matrix   <- matrix(NA, nrow = validation_number, ncol = n_coefs)
  colnames(coefs_matrix) <- coef_names
  colnames(var_matrix)   <- coef_names
  
  n_converged  <- 0                        # converged iterations collected
  n_attempts   <- 0                        # total attempts made
  max_attempts <- validation_number * 10   # stop if too many failures
  current_seed <- 1
  
  while (n_converged < validation_number && n_attempts < max_attempts) {
    
    n_attempts   <- n_attempts + 1
    set.seed(current_seed)
    current_seed <- current_seed + 1
    
    samp1 <- sample(1:nrow(data_realizations),
                    size = floor(2/3 * nrow(data_realizations)))
    samp0 <- sample(1:nrow(data_nonrealizations),
                    size = floor(2/3 * nrow(data_nonrealizations)))
    sample_data <- rbind(data_realizations[samp1, ],
                         data_nonrealizations[samp0, ])
    
    
    model <- glmer(formula_obj, family = binomial("logit"), data = sample_data,
                  control = glmerControl(optimizer = "bobyqa",
                                       optCtrl = list(maxfun = 2e5)))
    
    #check convergence
    if (!is.null(model@optinfo$conv$lme4$messages)) {
      n_converged <- n_converged + 1
      
      vec_coefs <- fixef(model)
      vec_var   <- diag(vcov(model))
      idx       <- match(names(vec_coefs), coef_names)
      idx_valid <- !is.na(idx)
      coefs_matrix[n_converged, idx[idx_valid]] <- vec_coefs[idx_valid]
      var_matrix[n_converged,   idx[idx_valid]] <- vec_var[idx_valid]
      
      cat("Attempt", n_attempts, "| Converged:", n_converged, "/",
          validation_number, "\n")
    } else {
      cat("Attempt", n_attempts, "| Did not converge — skipping\n")
    }
  }
  
  # report final status
  if (n_converged < validation_number) {
    warning(paste0("Only ", n_converged, " of ", validation_number,
                   " converged models collected after ", 
                   max_attempts, " attempts."))
  }
  
  # trim matrices to actual converged rows
  coefs_matrix <- coefs_matrix[1:n_converged, , drop = FALSE]
  var_matrix   <- var_matrix[1:n_converged,   , drop = FALSE]
  
  # King (2001) combination rule — use n_converged throughout
  mean_coef   <- colMeans(coefs_matrix, na.rm = TRUE)
  W           <- colMeans(var_matrix,   na.rm = TRUE)
  B           <- apply(coefs_matrix, 2, var, na.rm = TRUE)
  T_var       <- W + (1 + 1 / n_converged) * B   # n_converged not n
  se_combined <- sqrt(T_var)
  z_value     <- mean_coef / se_combined
  p_value     <- 2 * (1 - pnorm(abs(z_value)))
  stars <- ifelse(p_value < 0.001, "***",
           ifelse(p_value < 0.01,  "**",
           ifelse(p_value < 0.05,  "*",
           ifelse(p_value < 0.1,   ".", ""))))
  
  results <- data.frame(
    Variable     = coef_names,
    Coefficient  = round(mean_coef,   6),
    `Std. Error` = round(se_combined, 6),
    `z value`    = round(z_value,     3),
    `Pr(>|z|)`   = round(p_value,     4),
    Sig          = stars,
    check.names  = FALSE
  )
  
  return(list(
    results        = results,
    combined_coefs = mean_coef,
    combined_se    = se_combined,
    coefs_matrix   = coefs_matrix,
    var_matrix     = var_matrix,
    n_converged    = n_converged,
    n_attempts     = n_attempts
  ))
}
```

```{r cross-val-application}
#some models did not converge
val_num <- 5
cross_val <- cross_validation_fun_glmer(data = df_final_methods, dep_var = 'NVC2.1_VIOL', validation_number = val_num)
```

```{r h1-model-crossval-table, include = TRUE}
# Extract main model coefficients and SEs
main_coefs <- fixef(model1)
main_se    <- sqrt(diag(vcov(model1)))

# Cross-validated results
cv_coefs <- cross_val$combined_coefs
cv_se    <- cross_val$combined_se

var_labels <- c(
  "epr_IFI_l"         = "IFI",
  "epr_EFI_l"         = "EFI",
  "egip_participated_l" = 'Ethnic Group in Power Participation',
  "log_gdp_pcap_l"               = "ln(GDP per capita)",
  "pop_log_l"                    = "ln(Population)",
  "v2x_execorr_l"                = "Executive Corruption",
  "v2x_polyarchy_l"              = "Polyarchy",
  "log_v2regdur_l"               = "ln(Regime Duration)",
  "peace_years_l"                = "Peace Years"
)

# Combine into table
comparison_df <- data.frame(
  Variable        = names(main_coefs),
  `Main Model`    = sprintf("%.3f (%.3f)", main_coefs, main_se),
  `Cross-Validated` = sprintf("%.3f (%.3f)", cv_coefs, cv_se),
  check.names     = FALSE
) %>%
  filter(Variable != "(Intercept)") %>%
  mutate(Variable = dplyr::recode(Variable, !!!var_labels))  # apply labels

comparison_df %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Model 1: Main vs. Cross-Validated Coefficients",
    label    = "cv-comparison",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(
    general = "Intercept is not reported. Standard errors in parentheses. Cross-validated SE via King (2001) combination rule.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable    = TRUE 
  )
```

Finally, we cross-validate the results (`r val_num` cross-validations) to check whether the same results hold when models are estimated on different random subsamples. For this, we define the `cross_validation_fun` function^[Cross-validation subsampling causes convergence issues due to low sample sizes of the subsamples. To address this, we force the function to resample and reestimate the model until convergence. Thus, it can take more than one attempt to estimate one cross-validated subsample. It took `r cross_val$n_attempts` attempts to estimate `r val_num` cross-validations]. For this part, we did not calculate quantities of interest, but rather compare coefficient values and their significance, which will be enough to judge whether the results are stable or not. Table \@ref(tab:cv-comparison) shows the coefficients differences between the main and the cross-validated model. For both models, the IFI coefficient is negative and insignificant, though the coefficient absolute values differ. Moreover, one could notice a slight inflation of the standard errors, though not significantly changing the null result. Moreover, the reader should account for the sample size and respective test sample sizes that partly explain standard errors' inflation.

\section{Discussion}

\section{Conclusion}

\newpage

# References

<div id="refs"></div>

\newpage

# Appendix {-}
\appendix
\renewcommand{\thesection}{\Alph{section}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}

## Appendix A: AUC - ROC Curve {#appendix-a}

```{r h1-auc-roc-appendix, fig.cap="AUC-ROC Curve - Model 1", fig.align = 'center', out.width="85%", include = TRUE}
# Step 1: train-test split (same seed as confusion matrix section)
set.seed(666)

pred_test <- predict(model1_train, newdata = df_final_methods_test,
                     type = "response",
                     allow.new.levels = TRUE)

valid_idx    <- !is.na(pred_test) & !is.na(df_final_methods_test$NVC2.1_VIOL)
actual_test  <- df_final_methods_test$NVC2.1_VIOL[valid_idx]
probs_test   <- pred_test[valid_idx]

roc_obj <- roc(response  = actual_test,
               predictor = probs_test,
               smooth    = FALSE,      # kernel smoothing
               quiet     = TRUE)

roc_df <- data.frame(
  FPR = 1 - roc_obj$specificities,
  TPR = roc_obj$sensitivities
)

ggplot(roc_df, aes(x = FPR, y = TPR)) +
  geom_ribbon(aes(ymin = 0, ymax = TPR),
              fill = "#C0392B", alpha = 0.08) +
  geom_line(color = "#C0392B", linewidth = 0.9) +
  geom_abline(slope     = 1, intercept = 0,
              linetype  = "dashed", color = "grey60",
              linewidth = 0.5) +
  annotate("text",
           x     = 0.72, y = 0.25,
           label = paste0("AUC = ", round(auc(roc_obj), 3)),
           color = "#C0392B", size = 4) +
  scale_x_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  labs(
    x       = "False positive rate (1 - Specificity)",
    y       = "True positive rate (Sensitivity)"
  ) +
  theme_bw() +
  theme(
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(size = 8, color = "grey50", hjust = 1)
  )
```
\newpage

## Appendix B: Within-Between Random Effects Robustness Check {#appendix-b}

```{r h1-robustness-plot-appendix, fig.cap="Simulated predicted probabilities of a violent tactic choice over a range of IFI values - baseline vs. REWB", fig.align = 'center', out.width="85%", include = TRUE}
ova_egip_full_m1_appendix <- ova_egip_full_m1 %>% mutate(type = "RE")
  
rewb_model <- glmer(NVC2.1_VIOL ~
                      demeaned_IFI + mean_IFI +
                      demeaned_EFI + mean_EFI +
                      log_gdp_pcap_l + pop_log_l +
                      v2x_execorr_l + v2x_polyarchy_l +
                      log_v2regdur_l + peace_years_l +
                      (1 | cow),
                    data   = df_final_methods,
                    family = binomial("logit"))

ova_egip_full_rewb  <- sim_ova_range(rewb_model,        
                                 focal_var = "mean_IFI", 
                                 nsim = simulation_number,
                                 vcov_matrix = vcov(rewb_model))$result %>% mutate(type = "REWB")

ova_model1_robustness_appendix <- bind_rows(
  ova_egip_full_m1_appendix,
  ova_egip_full_rewb
) %>%
  mutate(type = factor(type, levels = c(
    "RE",
    "REWB"
  )))

rug_data <- bind_rows(
  df_final_methods %>% drop_na(epr_IFI_l) %>%
    mutate(type = "RE", rug_x = epr_IFI_l),
  df_final_methods %>% drop_na(mean_IFI) %>%
    mutate(type = "REWB", rug_x = mean_IFI)
) %>%
  mutate(type = factor(type, levels = c("RE", "REWB")))

ggplot(ova_model1_robustness_appendix,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2), 
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  geom_rug(
    data        = rug_data,
    aes(x       = rug_x),
    inherit.aes = FALSE,
    color       = "#C0392B", alpha = 0.3,
    sides       = "b"
    ) + 
  labs(x     = "IFI (0 = fully fragmented, 1 = fully concentrated)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = paste0(format(simulation_number), " simulations, 95% CI")) +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```

\newpage

## Appendix C: Campaigns excluded from the empirical analysis due to country coding issues {#appendix-c}
```{r lost-campaigns-appendix, include = TRUE}
lost_campaigns_table %>%
  arrange(country, year) %>%
  kbl(
    booktabs  = TRUE,
    longtable = TRUE,                # allows table to span multiple pages
    caption   = "Campaigns excluded from the analysis sample due to states panel coverage",
    label     = "lost-campaigns-appendix",
    col.names = c("COW code", "Year", "Country"),
    align     = c("c", "c", "l")
  ) %>%
  kable_styling(
    latex_options = c("striped", "repeat_header"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(
    general           = paste0(
      "N = ", nrow(lost_campaigns_table), " campaigns. ",
      "For the reasons os exclusion: see a footnote in the introduction"
    ),
    general_title     = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable    = TRUE
  )
```

