---
title: "**Guns or Gandhi: Ethnic Elite Composition and Dissident Tactics in Civil Conflicts**"
author: 
  - "Ruslan Guseinov"
  - "Maaike Arienne"
output:
    bookdown::pdf_document2:
    toc: true
    number_sections: false
  #pdf_document:
  #  toc: true
linestretch: 2
indent: true
fontsize: 12pt
geometry: left=2.5cm, right=2.5cm, top=2.5cm, bottom=2.5cm
date: "`r Sys.Date()`"
bibliography: references.bib
header-includes:
  - \setcounter{secnumdepth}{0}
  - \usepackage{amsthm}
  - \newtheorem{hyp}{Hypothesis}
abstract: |
  This paper is very cool blbablalblab
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,    # hide code from output
  warning = FALSE,    # suppress warnings
  message = FALSE,     # suppress package loading messages
  include = FALSE # hide code
)

p_needed <- c('tidyverse', 'mlogit', 'nnet', 'margins', 'marginaleffects', 'broom', 'summarytools', "MASS", 'kableExtra', 'stargazer', 'sandwich', 'sampleSelection', 'grateful', 'modelsummary', 'bookdown', 'patchwork', 'MASS', 'lme4')
packages <- rownames(installed.packages())
p_to_install <- p_needed[!(p_needed %in% packages)]
if (length(p_to_install) > 0) {
  install.packages(p_to_install) 
}
sapply(p_needed, require, character.only = TRUE)
options(warn = -1)

download.file(
  "https://www.zotero.org/styles/apa",
  destfile = "apa.csl"
)

stargazer_opt <- ifelse(knitr::is_latex_output(), "latex", "html")
```

```{r functions}
source('simulation_functions.R')
source('ova_sim_functions.R')
```

```{r load-data}
df_final_regr <- readRDS('datasources/df_essay.rds') #Ruslan's dataset
# RESTRICT THE DATASET TO ONLY ONSETS
# TEST THE HYPOTHESES ABOUT THE CHOICE OF METHODS
df_final_methods <- df_final_regr %>% 
  filter(onset_type %in% c('Nonviolent', 'Violent'))
```

**Keywords:** ethnic conflict, ethnic power relations, civil resistance, maximalist campaigns

\section{Introduction}

\section{Theory \& Hypotheses}

\begin{hyp}
 A higher count of ethnic groups in power increases the likelihood that dissident campaigns adopt violent tactics over nonviolent tactics.
\end{hyp}

\section{Data and Methods}
\subsection{Data Description}
Our study employs a panel dataset of `r nrow(df_final_methods)` so-called "maximalist campaigns", a concept firstly introduced by Chenoweth and Stephan in *Why Civil Resistance Works* book [@chenoweth2011civil]. Following their definition, a campaign is "a series of observable, continuous, purposive mass tactics or events in pursuit of a political objective’, with maximalist objectives encompassing regime change, self-determination, secession, and the expulsion of foreign occupations [@DVN/MHOXDV_2019]. We rely on Nonviolent and Violent Campaigns and Outcomes (NAVCO) 2.1 dataset, which records 384 maximalist campaigns from 1945 to 2013, extended by John Chin in his dissertation to `r nrow(df_final_methods)` campaigns - the version we use [@chin2017military].

The central conceptual distinction is between violent and nonviolent repertoires of contention. Following NAVCO coding rules, nonviolent tactics ‘do not directly threaten or harm the physical wellbeing of the regime, its agents, or its citizens’ and the violent repertoire ‘involves the use of force to physically harm or threaten to harm the opponents’ [@DVN/MHOXDV_2019]. The categories are mutually exclusive. A prominent example of a nonviolent campaign is the German 1989 *Die Wende* and the Cuban Revolution exemplifies a violent campaign. The violence/nonviolence distinction is well-established in the peace science literature and some significant findings have been made using NAVCO dataset [@celestino2013fresh; @manekin2022effective; @cunningham2013understanding], some papers also rely on John Chin's NAVCO 2.1 extension, which we use [@chin2023personalization]. Violence is operationally coded as a binary variable (1 = a campaign take a violent form, 0 = a campaign takes a nonviolent form).

Our key independent variables are drawn from a Ethnic Power Relations (EPR) Core Dataset [@vogt2015integrating]. EPR provides information on ethnic groups and their relative access to power. Inspired by Bivens and Thurber [-@Bivens26], we focus on the Ethnic Groups in Power (EGIP) and Marginalized Ethnic Groups (MEG^[We also refer to MEG as "excluded groups", these concepts are used synonymously]) counts. EGIP encompasses groups that exercise political power either via maintaining ethnic dominance or through power-sharing arrangements with junior partners, whilst MEG covers groups that are politically excluded from decision making, including both powerless and purposely discriminated groups. Due to theoretical expectations that a marginal effect of the ethnic groups count on the tactic choice diminishes at higher values, we apply the natural logarithm transformations to both EGIP and MEG counts in the empirical analysis.

To address endogeneity and block non-causal paths, we control for a standard set of confounders: GDP per capita in constant PPP dollars [@gapminderGDPCapitaConstant2024]; GDP annual growth rates; logged population size [@unitednationsWorldPopulationProspects2024]; youth bulge - the proportion of young population (15-29) to the total population size; executive corruption scores [@coppedgeVDemCodebookV142024]; polyarchy (electoral democracy) scores [@coppedgeVDemCodebookV142024]; regime duration in years [@coppedgeVDemCodebookV142024]; peace years (years elapsed since the last maximalist campaign in each country); EGIP population share to the total population, to account for a possible confounding effect between EGIP count and a tactic choice; logged Marginalized Ethnic Groups to accout for a possible confounding as well. To address reversed causality, all the independent variables are one-year lagged^[Even though we find it unconvincing that a choice of one tactic over the other could anyhow affect the number of EGIP and MEG].

\subsection*{Descriptive statistics}
```{r descriptive-stats, echo=FALSE, include = TRUE}
datasummary(
  (`EGIP Groups (count)`        = epr_egip_groups_count_l) +
  (`Excluded Groups (count)`    = epr_excl_groups_count_l) +
  (`EGIP Population Share`      = epr_egippop_l) +
  (`GDP per capita (log)`       = log_gdp_pcap_l) +
  (`Population (log)`           = pop_log_l) +
  (`Youth Bulge`                = youthbulge_l) +
  (`GDP Growth`                 = gdp_growth_l) +
  (`Executive Corruption`       = v2x_execorr_l) +
  (`Polyarchy`                  = v2x_polyarchy_l) +
  (`Regime Duration (log)`      = log_v2regdur_l) +
  (`Peace Years`                = peace_years_l) ~
  Mean + SD + Min + Max + N,
  data  = df_final_methods,
  title = "Descriptive Statistics - Campaign Onset Sample",
  label = "descriptive-stats",
  output = "kableExtra"         
) %>% kableExtra::kable_styling(latex_options = c("hold_position"))
```

Table \@ref(tab:descriptive-stats) presents descriptive statistics for all variables included in the analysis. Data missingness is not a significant issue, though for all campaigns in Yemen there is missing data on ethnic groups. Figure \@ref(fig:epr-group-count) shows the distribution of logged ethnic groups counts with respect to the tactic choice. At low EGIP and MEG counts, nonviolent tactics are more preferred and with the growing ethnic groups numbers the violent tactic is prevalent. The modal logged count is 1 - approximately `r round(exp(1), 2)` groups - for both EGIP and MEG, regardless of tactic. At higher values of both EGIP and MEG, campaigns are much more rare. This visualization shows some degree of variation 1) between violent and nonviolent choice (which is the main focus of the essay) and 2) between EGIP and MEG (which is less interesting for us).

```{r epr-group-count, fig.cap="Ethnic Group Counts by Repertoire Choice", fig.align = 'center', out.width="85%", include = TRUE}
df_final_methods %>%
  dplyr::select(onset_type,
         `EGIP Groups (logged)`     = log_epr_egip_groups_count_l,
         `Excluded Groups (logged)` = log_epr_excl_groups_count_l) %>%
  pivot_longer(-onset_type, names_to = "variable", values_to = "count") %>%
  ggplot(aes(x = count, fill = onset_type)) +
  geom_histogram(binwidth = 1, position = "dodge", color = "white") +
  scale_fill_manual(values = c("Nonviolent" = "#2C86C7",
                                "Violent"    = "#C0392B")) +
  facet_wrap(~variable) +
  labs(x = "Logged group count", y = "Count", fill = NULL) +
  theme_bw() +
  theme(legend.position = "bottom")
```

\subsection*{Modeling approach}

Given a binary dependent variable and a balanced sample -`r nrow(df_final_methods[df_final_methods$NVC2.1_VIOL == 1,])` of violent and `r nrow(df_final_methods[df_final_methods$NVC2.1_VIOL == 0,])` of nonviolent onsets - we estimate a logit model. As we do not predict the duration of events, there is no need for advanced binary choice models such as a complementary log-log. We also include a country fixed effect to control for the time-invariant unobserved factors (such as culture or geography).

For Hypothesis 1, our model specification is:
$$
\begin{aligned}
& \text{violent}_{it} \sim \text{Bernoulli}(\pi_{it}) \\[0.3em]
& \log\left(\frac{\pi_{it}}{1-\pi_{it}}\right) = \beta_0 
+ \beta_1 \ln(\text{EGIP})_{it-1} 
+ \beta_2 \ln(\text{MEG})_{it-1} 
+ \mathbf{X}'_{it-1}\boldsymbol{\gamma} + \alpha_i
&
\end{aligned}
$$
where $\mathbf{X}_{i}$ is the vector of covariates and $\alpha_i$ is a country fixed effect.

For uncertainty calculations, we follow a King's simulation approach [@king2000making]. This approach is also well-established in peace science literature [@butcher2016manufacturing]. Instead of reporting coefficients and their significance, we calculate predicted probabilities as the main quantity of interest via an observed value approach (OVA), meaning we simulate the predicted probabilities for all observations and then calculate the mean of them^[We report coefficients only in the cross-validation robustness check]. For the essay, we define `sim_ova_range` function for the OVA simulations, allowing to calculate predicted probabilities in just four lines of code. It also allows for different CI levels, seeds, number of values and focal variable ranges at the researcher's discretion.

\subsection{Results}

```{r}
re_model <- glmer(NVC2.1_VIOL ~ 
                    epr_IFI_l + epr_EFI_l + log_epr_egippop_l + 
                    log_gdp_pcap_l + pop_log_l + youthbulge_l + gdp_growth_l +
                    v2x_execorr_l + v2x_polyarchy_l +
                    log_v2regdur_l + peace_years_l + 
                    (1 | cow),        # random intercept by country
                  data = df_final_methods,
                  family = binomial(link = "logit"))

ova_egip_full_m1  <- sim_ova_range(re_model,        
                                 focal_var = "epr_IFI_l", 
                                 nsim = 5000,
                                 vcov_matrix = vcov(re_model))$result %>% mutate(type = "Full Sample")

ggplot(ova_egip_full_m1,
       aes(x = focal_value, y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2), 
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  labs(x     = "Included Fractionalization Index", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = "1000 simulations, 95% CI") +
  theme_bw() +
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```


```{r h1-modeling}
model1 <- glm(NVC2.1_VIOL ~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l + log_epr_egippop_l + 
   log_gdp_pcap_l + pop_log_l + youthbulge_l + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l +
    log_v2regdur_l + peace_years_l + as.factor(cow),
    family = binomial(link = "logit"),
    data = df_final_methods)
```

```{r h1-qoisim}
ova_egip_full_m1  <- sim_ova_range(model1,        
                                 focal_var = "log_epr_egip_groups_count_l", 
                                 nsim = 1000,
                                 vcov_matrix = vcov(model1))$result %>% mutate(type = "Full Sample")
```

```{r h1-plot, fig.cap="Simulated predicted probabilities of a violent tactic choice over a range of ethnic groups in power counts", fig.align = 'center', out.width="85%", include = TRUE}
ggplot(ova_egip_full_m1,
       aes(x = exp(focal_value), y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2), 
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  scale_x_continuous(limits = c(NA, 7), breaks = 0:7) +
  labs(x     = "Ethnic Groups in Power (count)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = "1000 simulations, 95% CI") +
  theme_bw() +
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```
To test the first hypothesis, we estimate a logit model of the violent tactic choice, according to the H1 model specification. Figure \@ref(fig:h1-plot) shows the simulated predicted probabilities of a violent tactic choice over a range of observed EGIP counts, with focal values exponentiated for a better interpretation. The x-axis is truncated at seven ethnic groups counts to filter out sparse data and unrealistic extrapolations. Despite a downward trend - from `r round(ova_egip_full_m1$mean[1], 2)` at one group to `r round(ova_egip_full_m1$mean[50], 2)` at seven groups - confidence intervals that encompass 0.5 do not allow us to claim that with the increasing number of ethic groups in power the rebels would prefer violence over nonviolence. Instead of zero probability, our benchmark probability is approx. `r round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2)` - the sample proportion of violent campaigns in the whole sample.


```{r h1-train-test sampling, include = TRUE}
set.seed(666)  # same seed as simulations for reproducibility

n_total <- nrow(df_final_methods) #define the number of observations
train_idx <- sample(1:n_total, size = floor(2/3 * n_total), replace = FALSE) #slice two thirds into the train set
test_idx  <- setdiff(1:n_total, train_idx) #all other observations into the test

df_final_methods_train <- df_final_methods[train_idx, ] # train dataset

model1_train <- glm(NVC2.1_VIOL ~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l + log_epr_egippop_l + 
   log_gdp_pcap_l + pop_log_l + youthbulge_l + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l +
    log_v2regdur_l + peace_years_l + as.factor(cow),
    family = binomial(link = "logit"),
    data = df_final_methods_train) #estimate the model on a train set

trained_cows <- model1_train$xlevels$`as.factor(cow)` #what country FEs are in the estimated model?
df_final_methods_test <- df_final_methods[test_idx, ] %>%
  filter(as.character(cow) %in% trained_cows) %>% #test is only with the trained country FEs
  drop_na(log_epr_egip_groups_count_l, log_epr_excl_groups_count_l,
          log_epr_egippop_l, log_gdp_pcap_l, pop_log_l, youthbulge_l,
          gdp_growth_l, v2x_execorr_l, v2x_polyarchy_l,
          log_v2regdur_l, peace_years_l)

pred_test_prob <- predict(model1_train, newdata = df_final_methods_test, type = "response")  #predict on the test dataset
actual_test <- ifelse(df_final_methods_test$NVC2.1_VIOL == 1,
                      "Violent", "Nonviolent")
pred_test   <- ifelse(pred_test_prob >= 0.5, "Violent", "Nonviolent")

tp <- sum(pred_test == "Violent"    & actual_test == "Violent")
tn <- sum(pred_test == "Nonviolent" & actual_test == "Nonviolent")
fp <- sum(pred_test == "Violent"    & actual_test == "Nonviolent")
fn <- sum(pred_test == "Nonviolent" & actual_test == "Violent")

accuracy  <- (tp + tn) / (tp + tn + fp + fn)
precision <- tp / (tp + fp)
recall    <- tp / (tp + fn)   
f1        <- 2 * (precision * recall) / (precision + recall)

conf_df <- data.frame(
  ` `            = c("Predicted Nonviolent", "Predicted Violent"),
  `Actual Nonviolent` = c(as.character(tn), as.character(fp)),
  `Actual Violent`    = c(as.character(fn), as.character(tp)),
  check.names = FALSE
)

metrics_df <- data.frame(
  ` `                 = c("Accuracy", "Precision", 
                           "Recall",  "F1 Score"),
  `Actual Nonviolent` = c(sprintf("%.3f", accuracy), 
                           sprintf("%.3f", precision),
                           sprintf("%.3f", recall), 
                           sprintf("%.3f", f1)),
  `Actual Violent`    = c("", "", "", ""),
  check.names = FALSE
)

bind_rows(conf_df, metrics_df) %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Confusion Matrix and Predictive Performance — Model 1",
    label    = "conf-matrix-h1",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  row_spec(2, extra_css = "border-bottom: 2px solid black;") %>%  # separator line
  pack_rows("Confusion Matrix", 1, 2) %>%
  pack_rows("Performance Metrics", 3, 6)
```

Despite an insignificant effect of EGIP on a violent choice, the model as a whole demonstrates a strong predictive performance. We divided the dataset into a train and test sample, estimated a logit model on the train sample and predicted a tactic choice on the test set. Table \@ref(tab:conf-matrix-h1) reports the confusion matrix and performance metrics. With an accuracy of `r round(accuracy, 2)` and F1 score of `r round(f1, 2)`, our model in general predicts tactic choice well above a coin toss, predicting nonviolence a bit better than the violent choice. An important implication here is that the model as a whole (including the specification itself, a set of covariates and a country fixed effect) captures meaningful variation in the outcome, even though the EGIP does not significantly predict a violent choice.

\subsection{Robustness checks}
We conduct three robustness checks. First, given that civil conflicts and a tactic choice are regionally clustered, we estimate an additional model on the Middle East and North Africa (MENA) and Sub-Saharan Africa countries subsample, where violent campaigns are more expected than in other regions. For the full sample, there are `r round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods) * 100, 1)`% of violent campaigns, whilst for MENA \& Sub-Saharan Africa there are `r round(sum(df_final_methods[df_final_methods$region %in% c('Middle East & North Africa', 'Sub-Saharan Africa'),]$NVC2.1_VIOL) / nrow(df_final_methods[df_final_methods$region %in% c('Middle East & North Africa', 'Sub-Saharan Africa'),]) * 100, 1)`% of them. Second, territorial campaigns could be more ethnicity-driven and secessionist civil wars along ethnic lines are widely researched in the academic literature [@collier2006political; @denny2014ethnicity; @toft2012self]. For this reason, we estimated three separate models 1) on the MENA & Sub-Saharan Africa subsample, 2) excluding territorial campaigns and 3) both on the MENA & Sub-Saharan Africa sub-sample with excluded territorial campaigns.

```{r h1-robustness-modelling}
model1_africa <- glm(NVC2.1_VIOL ~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l + log_epr_egippop_l + 
    log_gdp_pcap_l + pop_log_l + youthbulge_l  + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l +
    log_v2regdur_l +  peace_years_l + as.factor(cow),
    family = binomial(link = "logit"),
    data = df_final_methods %>% filter(region %in% c('Middle East & North Africa', 'Sub-Saharan Africa')))

model1_noterr <- glm(NVC2.1_VIOL ~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l + 
   log_gdp_pcap_l + pop_log_l + youthbulge_l + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l +
    log_v2regdur_l + peace_years_l + I(peace_years_l^2) + I(peace_years_l^3),
    family = binomial(link = "logit"),
    data = df_final_methods %>% filter(NVC2.1_territorial == 0))

model1_africa_noterr <- glm(NVC2.1_VIOL ~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l + 
    log_gdp_pcap_l + pop_log_l + youthbulge_l  + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l +
    log_v2regdur_l +  peace_years_l + I(peace_years_l^2) + I(peace_years_l^3),
    family = binomial(link = "logit"),
    data = df_final_methods %>% filter(region %in% c('Middle East & North Africa', 'Sub-Saharan Africa')) %>% filter(NVC2.1_territorial == 0))
```

```{r h1-robustness-qoisim}
ova_egip_africa_m1  <- sim_ova_range(model1_africa,        
                                 focal_var = "log_epr_egip_groups_count_l", 
                                 nsim = 1000,
                                 vcov_matrix = vcov(model1_africa))$result %>% mutate(type = "MENA & Sub-Saharan Africa")
ova_egip_noterr_m1  <- sim_ova_range(model1_noterr,        
                                 focal_var = "log_epr_egip_groups_count_l", 
                                 nsim = 1000,
                                 vcov_matrix = vcov(model1_noterr))$result %>% mutate(type = "Excl. territorial campaigns")
ova_egip_africa_noterr_m1  <- sim_ova_range(model1_africa_noterr,        
                                 focal_var = "log_epr_egip_groups_count_l", 
                                 nsim = 1000,
                                 vcov_matrix = vcov(model1_africa_noterr))$result %>% mutate(type = "MENA & Sub-Saharan Africa & Excl. terr.")

ova_model1_robustness <- bind_rows(
  ova_egip_full_m1,
  ova_egip_africa_m1,
  ova_egip_noterr_m1,
  ova_egip_africa_noterr_m1
) %>%
  mutate(type = factor(type, levels = c(
    "Full Sample",
    "MENA & Sub-Saharan Africa",
    "Excl. territorial campaigns",
    "MENA & Sub-Saharan Africa & Excl. terr."
  )))
```

```{r h1-robustness-plot, fig.cap="Simulated predicted probabilities of a violent tactic choice over a range of ethnic groups in power counts - robustness checks results", fig.align = 'center', out.width="85%", include = TRUE}
ggplot(ova_model1_robustness,
       aes(x = exp(focal_value), y = mean)) +
  geom_line(linewidth = 0.9, color = "#C0392B") +
  geom_hline(yintercept = round(sum(df_final_methods$NVC2.1_VIOL) / nrow(df_final_methods), 2), 
             linetype = 'dashed', color = 'grey', alpha = 0.8) + 
  geom_ribbon(aes(ymin = lower, ymax = upper), 
              alpha = 0.15, fill = "#C0392B", color = NA) +
  scale_x_continuous(limits = c(NA, 7), breaks = 0:7) +
  labs(x     = "Ethnic Groups in Power (count)", 
       y     = "Predicted Probability\n(Violent vs. Nonviolent)", 
       caption = "1000 simulations, 95% CI") +
  theme_bw() +
  facet_wrap(~type) + 
  theme(
    axis.title.y     = element_text(margin = margin(r = 10)),
    panel.grid.minor = element_blank(),
    plot.caption     = element_text(hjust = 1, color = "grey50")
  )
```

We calculate simulated predicted probabilities over the range of EGIP counts for all three robustness check models and summarize the findings on one plot. Figure \@ref(fig:h1-robustness-plot) shows the results along three robustness specifications, including the full-sample baseline The only partially significant finding is in the model which excludes territorial campaigns (this sub-sample mostly includes anti-regime campaigns). For a low count of ethnic groups in power, campaigns that exclude territorial goals/claims have a higher chance of being nonviolent and the rebels push towards a higher likelihood of violence with the growing number of ethnic groups in power. This (though cautious) finding can be explained by the logic of mass anti-regime mobilization, where a more fragmented ethnic elite decreases the benefits of nonviolent resistance and pushes the rebels toward violent repertoires of contention. For the MENA \& Sub-Saharan Africa subsample (both with and without territorial campaigns) there is no systematic difference between violence and nonviolence across the growing number of EGIP. In general, our null result is robust.

```{r cross-validation-function}
cross_validation_fun <- function(validation_number = 10,
                                   data,
                                   dep_var) {
  n <- validation_number # set the number of validations
  seed_vectors <- seq_len(n) # a set of seeds (increasing number of validations)

  # model specification
  formula_obj <- as.formula(
    paste(dep_var, "~ log_epr_egip_groups_count_l + log_epr_excl_groups_count_l +
    log_epr_egippop_l + log_gdp_pcap_l + pop_log_l + youthbulge_l + gdp_growth_l +
    v2x_execorr_l + v2x_polyarchy_l + log_v2regdur_l + peace_years_l + as.factor(cow)")
  )

  # balance the sample by realizations and non-realizations
  data_realizations    <- data[data[[dep_var]] == 1, ]
  data_nonrealizations <- data[data[[dep_var]] == 0, ]

  # Get substantive coefficient names — drop country FE
  test_fit   <- glm(formula_obj, family = binomial("logit"), data = data)
  all_names  <- names(coef(test_fit))
  keep_idx   <- !grepl("as.factor\\(cow\\)", all_names)  # exclude FE
  coef_names <- all_names[keep_idx]
  n_coefs    <- length(coef_names)

  coefs_matrix <- matrix(NA, nrow = n, ncol = n_coefs)
  var_matrix   <- matrix(NA, nrow = n, ncol = n_coefs)
  colnames(coefs_matrix) <- coef_names
  colnames(var_matrix)   <- coef_names

  # loop through validations
  for (i in 1:n) {
    set.seed(seed_vectors[i]) # set a seed

    samp1 <- sample(1:nrow(data_realizations),
      size = floor(2 / 3 * nrow(data_realizations))
    ) # sample realizations
    samp0 <- sample(1:nrow(data_nonrealizations),
      size = floor(2 / 3 * nrow(data_nonrealizations))
    ) # sample non-realizations

    sample_data <- rbind(
      data_realizations[samp1, ],
      data_nonrealizations[samp0, ]
    ) # combine the data

    model  <- glm(formula_obj, family = binomial("logit"), data = sample_data)
    
    vec_coefs <- coef(model)
    vec_var   <- diag(vcov(model))

    # Keep only substantive coefficients — drop FE
    keep      <- !grepl("as.factor\\(cow\\)", names(vec_coefs))
    vec_coefs <- vec_coefs[keep]
    vec_var   <- vec_var[keep]

    # Align by name
    idx <- match(names(vec_coefs), coef_names)
    idx_valid <- !is.na(idx)
    coefs_matrix[i, idx[idx_valid]] <- vec_coefs[idx_valid]
    var_matrix[i,   idx[idx_valid]] <- vec_var[idx_valid]
  }

  # Apply the King (2001) rule

  # combined coefficients
  mean_coef <- colMeans(coefs_matrix, na.rm = TRUE)
  # within-sample variance
  W <- colMeans(var_matrix, na.rm = TRUE)
  # between-sample variance
  B <- apply(coefs_matrix, 2, var, na.rm = TRUE)
  # total variance
  T_var <- W + (1 + 1 / n) * B
  # combined standard errors
  se_combined <- sqrt(T_var)

  # inference
  z_value <- mean_coef / se_combined
  p_value <- 2 * (1 - pnorm(abs(z_value)))

  stars <- ifelse(p_value < 0.001, "***",
    ifelse(p_value < 0.01, "**",
      ifelse(p_value < 0.05, "*",
        ifelse(p_value < 0.1, ".", "")
      )
    )
  )

  results <- data.frame(
    Variable = coef_names,
    Coefficient = round(mean_coef, 6),
    `Std. Error` = round(se_combined, 6),
    `z value` = round(z_value, 3),
    `Pr(>|z|)` = round(p_value, 4),
    Sig = stars,
    check.names = FALSE
  )

  # return a list with the final results, coefficients and ses separately
  return(list(
    results = results,
    combined_coefs = mean_coef,
    combined_se = se_combined,
    coefs_matrix = coefs_matrix,
    var_matrix = var_matrix
  ))
}
```

```{r cross-val-application}
cross_val <- cross_validation_fun(data = df_final_methods, dep_var = 'NVC2.1_VIOL', validation_number = 10)
```

```{r h1-model-crossval-table, include = TRUE}
# Extract main model coefficients and SEs (no FE)
main_coefs <- coef(model1)
main_se    <- sqrt(diag(vcov(model1)))

# Keep only substantive coefficients
keep       <- !grepl("as.factor\\(cow\\)", names(main_coefs))
main_coefs <- main_coefs[keep]
main_se    <- main_se[keep]

# Cross-validated results
cv_coefs <- cross_val$combined_coefs
cv_se    <- cross_val$combined_se

var_labels <- c(
  "log_epr_egip_groups_count_l"  = "ln(EGIP Groups)",
  "log_epr_excl_groups_count_l"  = "ln(Excluded Groups)",
  "log_epr_egippop_l"            = "EGIP Population Share (ln)",
  "log_gdp_pcap_l"               = "ln(GDP per capita)",
  "pop_log_l"                    = "ln(Population)",
  "youthbulge_l"                 = "Youth Bulge",
  "gdp_growth_l"                 = "GDP Growth",
  "v2x_execorr_l"                = "Executive Corruption",
  "v2x_polyarchy_l"              = "Polyarchy",
  "log_v2regdur_l"               = "ln(Regime Duration)",
  "peace_years_l"                = "Peace Years"
)

# Combine into table
comparison_df <- data.frame(
  Variable        = names(main_coefs),
  `Main Model`    = sprintf("%.3f (%.3f)", main_coefs, main_se),
  `Cross-Validated` = sprintf("%.3f (%.3f)", cv_coefs, cv_se),
  check.names     = FALSE
) %>%
  filter(Variable != "(Intercept)") %>%
  mutate(Variable = dplyr::recode(Variable, !!!var_labels))  # apply labels

comparison_df %>%
  kbl(
    booktabs = TRUE,
    align    = c("l", "c", "c"),
    caption  = "Model 1: Main vs. Cross-Validated Coefficients",
    label    = "cv-comparison",
    escape   = FALSE
  ) %>%
  kable_styling(
    latex_options = c("hold_position"),
    full_width    = FALSE
  ) %>%
  row_spec(0, bold = TRUE) %>%
  footnote(
    general = "Intercept is not reported. Standard errors in parentheses. Cross-validated SE via King (2001) combination rule.",
    general_title = "Note:",
    footnote_as_chunk = TRUE,
    threeparttable    = TRUE 
  )
```

Finally, we cross-validate the results to check whether the same results hold when models are estimated on different random subsamples. For this, we define the `cross_validation_fun` function. For this part, we did not calculate quantities of interest, but rather compare coefficient values and their significance, which will be enough to judge whether the results are stable or not. Table \@ref(tab:cv-comparison) shows the coefficients differences between the main and the cross-validated model. For both models, the logged EGIP count coefficient is negative and insignificant, though the coefficient absolute values differ. Moreover, one could notice the inflation of the standard errors, suggesting the results are not as robust as expected. However, the reader should account for the sample size and respective test sample sizes that partly explain standard errors' inflation.

\section{Discussion}

\section{Conclusion}

# References

# Appendix {-}
\appendix
\renewcommand{\thesection}{\Alph{section}}

## A. First Appendix Section {-}

Content here...

## B. Second Appendix Section {-}

Content here...

